---
title: "Preparing past climate forcings for the HIMI model, based on Phoebe and Kieran's tutorial"
author: "Annabelle Erskine (edited from Phoebe Woodworth-Jefcoats & Kieran Murphy)"
date: "2023-10-04"
output: html_document
---
#to do- copy code from desktop version and tidy (annabelle)
```{r setup, include=FALSE}
library(tidyverse)
library(reshape2)
library(readr)
```

Extracted data (ISIMIP historical observed climate output,, ) from http://portal.sf.utas.edu.au/thredds/catalog/gem/fishmip/ISIMIP3a/InputData/climate/ocean/obsclim/regional/monthly/historical/GFDL-MOM6-COBALT2/catalog.html

HIMI model
15arcmin

## Model domain

### Check model domain area

Need a file that has the area_m2 info.
# Load total carbon data for picoplankton and reformat dataframe

```{r}
# Load total carbon data for picoplankton and reformat dataframe
df_diat_raw <- read_csv("gfdl-mom6-cobalt2_obsclim_phydiat-vint_15arcmin_himi_monthly_1961_2010.csv")
```

```{r}
df_area <- df_diat_raw %>% 
  select(area_m2)
  
# Hawaiian_Longline_area <- df_diat_raw %>%
#   select(c(lat,lon,area_m2)) %>%  
#   mutate(total_area = sum(area_m2))
# 
# head(Hawaiian_Longline_area)
# dim(Hawaiian_Longline_area)

HIMI_area <- df_diat_raw %>%
  select(c(lat,lon,area_m2)) %>%
  mutate(total_area = sum(area_m2))

head(HIMI_area)
dim(HIMI_area)
summary(HIMI_area)

dim(df_area)
summary(df_area)
```

#below is the hawaiian longline values
<!-- Each 1 degree grid cell -->   $9540434011 m2$ -->

<!-- Total model domain -->    $1.95e+13 m2$ -->

#AE: below : HIMI (is this right??) Take median value?

Each 1 degree grid cell -->   $7.864e+09   m2$ (AE: take median or mean value of area_m2?)

Total model domain -->    $1.719508e+12	 m2$

# Prepare plankton forcings

Need to do diatoms first to extract area vector
## Diatoms
```{r}
# Load total carbon data for picoplankton and reformat dataframe
# df_diat_raw <- read.csv("Plankton_Forcings/gfdl-mom6-cobalt2_obsclim_phydiat-vint_60arcmin_Hawaiian-Longline_monthly_1961_2010.csv") 

# AE Q: why are we doing this again, isn't this a repeat? ^^^

dim(df_diat_raw)
glimpse(df_diat_raw)
head(df_diat_raw)
```

```{r}
df_diat_long <- df_diat_raw %>%
  gather(Date, mol_C_m2,Jan_1961:Dec_2010) %>% # Convert from wide format to long
  mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
  mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
  mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
  mutate(date = parse_date_time(Date, orders = "my"))

df_area <- df_diat_long$area_m2
df_date <- df_diat_long$date

df_diat_long <- df_diat_long %>% # Create a tidy date variable
  group_by(date) %>% # group by month
  summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month

head(df_diat_long)
dim(df_diat_long)
```

## Picoplankton

```{r}
# Load total carbon data for picoplankton and reformat dataframe
# df_pico_raw <- read.csv("Plankton_Forcings/gfdl-mom6-cobalt2_obsclim_phypico-vint_15arcmin_Hawaiian-Longline_monthly_1961_2010.csv")

df_pico_raw <- read.csv("gfdl-mom6-cobalt2_obsclim_intppdiat_15arcmin_himi_monthly_1961_2010.csv")

dim(df_pico_raw)
glimpse(df_pico_raw)
head(df_pico_raw)
```


```{r}
# df_pico_long <- df_pico_raw %>%
#   select(!c(lat,lon)) %>% # remove lat, lon
#   melt() %>% # melt data from a wide array to long format
#   mutate(C_g_m2 = value * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = C_g_m2*9540434011) %>% # Calculate total g C per grid cell by multiplying by area in m^2 per grid cell
#   group_by(variable) %>% # group by month
#   summarise(total_C_g = sum(C_g)) %>%  # sum carbon so it is total carbon for whole model domain per month
#   mutate(date = seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")) %>% # create a tidy date variable
#   select(date, total_C_g) # select final variables


# df_pico_long <- df_pico_raw %>%
#   gather(Date, mol_C_m2, X1961.01.01.00.00.00:X2010.12.01.00.00.00) %>% # Convert from wide format to long
#   mutate(area_m2 = df_area) %>% 
#   mutate(date = df_date) %>% 
#   mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
#   mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
#   # mutate(date = parse_date_time(Date, orders = "my")) %>% # Create a tidy date variable
#   group_by(date) %>% # group by month
#   summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month

#AE what is going on above with the date range and decimals? Rewriting as above
df_pico_long <- df_pico_raw %>%
  gather(Date, mol_C_m2, Jan_1961:Dec_2010) %>% # Convert from wide format to long
  mutate(area_m2 = df_area) %>%
  mutate(date = df_date) %>%
  mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
  mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
  mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
  group_by(date) %>% # group by month
  summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month
```


*Important*
Need to update so the area is used grid cell by grid cell
AE: what is this part?

```{r}
# df_diat_long <- df_diat_raw %>%
#   select(!c(lat,lon,area_m2)) %>% # remove lat, lon, area_m2
#   melt() %>% # melt data from a wide array to long format
#   mutate(C_g_m2 = value * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = C_g_m2*9540434011) %>% # Calculate total g C per grid cell by multiplying by area in m^2 per grid cell
#   group_by(variable) %>% # group by month
#   summarise(total_C_g = sum(C_g)) %>%  # sum carbon so it is total carbon for whole model domain per month
#   mutate(date = seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")) %>% # create a tidy date variable
#   select(date, total_C_g) # select final variables
# 
# dim(df_diat_long)
# head(df_diat_long)
```

## Diazotrophs
```{r}
# Load total carbon data for picoplankton and reformat dataframe
# df_diaz_raw <- read.csv("Plankton_Forcings/gfdl-mom6-cobalt2_obsclim_phydiaz-vint_60arcmin_Hawaiian-Longline_monthly_1961_2010.csv")

df_diaz_raw <- read.csv("gfdl-mom6-cobalt2_obsclim_phydiaz-vint_15arcmin_HIMI_monthly_1961_2010.csv")

dim(df_diaz_raw)
glimpse(df_diaz_raw)
head(df_diaz_raw)
#AE funny date names here
```

```{r}
# df_diaz_long <- df_diaz_raw %>%
#   select(!c(lat,lon)) %>% # remove lat, lon
#   melt() %>% # melt data from a wide array to long format
#   mutate(C_g_m2 = value * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = C_g_m2*9540434011) %>% # Calculate total g C per grid cell by multiplying by area in m^2 per grid cell
#   group_by(variable) %>% # group by month
#   summarise(total_C_g = sum(C_g)) %>%  # sum carbon so it is total carbon for whole model domain per month
#   mutate(date = seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")) %>% # create a tidy date variable
#   select(date, total_C_g) # select final variables

df_diaz_long <- df_diaz_raw %>%
  gather(Date, mol_C_m2, X1961.01.01.00.00.00:X2010.12.01.00.00.00) %>% # Convert from wide format to long
  mutate(area_m2 = df_area) %>% 
  mutate(date = df_date) %>% 
  mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
  mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
  mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
  # mutate(date = parse_date_time(Date, orders = "my")) %>% # Create a tidy date variable
  group_by(date) %>% # group by month
  summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month
```

## Microzooplankton
```{r}
# Load total carbon data for picoplankton and reformat dataframe
# df_zmicro_raw <- read.csv("Plankton_Forcings/gfdl-mom6-cobalt2_obsclim_zmicro-vint_60arcmin_Hawaiian-Longline_monthly_1961_2010.csv")

df_zmicro_raw <- read.csv("gfdl-mom6-cobalt2_obsclim_zmicro-vint_15arcmin_himi_monthly_1961_2010.csv")

dim(df_zmicro_raw)

glimpse(df_zmicro_raw)
head(df_zmicro_raw)
```

```{r}
# df_zmicro_long <- df_zmicro_raw %>%
#   select(!c(lat,lon, area_m2)) %>% # remove lat, lon, area_m2
#   melt() %>% # melt data from a wide array to long format
#   mutate(C_g_m2 = value * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = C_g_m2*9540434011) %>% # Calculate total g C per grid cell by multiplying by area in m^2 per grid cell
#   group_by(variable) %>% # group by month
#   summarise(total_C_g = sum(C_g)) %>%  # sum carbon so it is total carbon for whole model domain per month
#   mutate(date = seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")) %>% # create a tidy date variable
#   select(date, total_C_g) # select final variables

df_zmicro_long <- df_zmicro_raw %>%
  gather(Date, mol_C_m2,Jan_1961:Dec_2010) %>% # Convert from wide format to long
  mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
  mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
  mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
  mutate(date = parse_date_time(Date, orders = "my")) %>% # Create a tidy date variable
  group_by(date) %>% # group by month
  summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month
```

## Mesozooplankton
```{r}
# Load total carbon data for picoplankton and reformat dataframe
# df_zmeso_raw <- read.csv("Plankton_Forcings/gfdl-mom6-cobalt2_obsclim_zmeso-vint_60arcmin_Hawaiian-Longline_monthly_1961_2010.csv")

df_zmeso_raw <- read_csv("gfdl-mom6-cobalt2_obsclim_zmeso-vint_15arcmin_himi_monthly_1961_2010.csv")
  
dim(df_zmeso_raw)
# glimpse(df_zmeso_raw)
# head(df_zmeso_raw)
```


```{r}
# df_zmeso_long <- df_zmeso_raw %>%
#   select(!c(lat,lon, area_m2)) %>% # remove lat, lon, area_m2
#   melt() %>% # melt data from a wide array to long format
#   mutate(C_g_m2 = value * 12.001) %>% # Convert from mol C m^2 to g C m^2
#   mutate(C_g = C_g_m2*9540434011) %>% # Calculate total g C per grid cell by multiplying by area in m^2 per grid cell
#   group_by(variable) %>% # group by month
#   summarise(total_C_g = sum(C_g)) %>%  # sum carbon so it is total carbon for whole model domain per month
#   mutate(date = seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")) %>% # create a tidy date variable
#   select(date, total_C_g) # select final variables

df_zmeso_long <- df_zmeso_raw %>%
  gather(Date, mol_C_m2,Jan_1961:Dec_2010) %>% # Convert from wide format to long
  mutate(C_g_m2 = mol_C_m2 * 12.001) %>% # Convert from mol C m^2 to g C m^2
  mutate(C_g = mol_C_m2*area_m2) %>% # g C per grid cell by multiplying by area in m^2 per grid cell
  mutate(C_gww = C_g * 10) %>% # *10 to get grams wet weight
  mutate(date = parse_date_time(Date, orders = "my")) %>% # Create a tidy date variable
  group_by(date) %>% # group by month
  summarise(total_C_gww = sum(C_gww)) # sum carbon so it is total carbon for whole model domain per month
```



## Create size mid points in grams wet weight (gww)

```{r}
# Create variables for referencing the size class mid points, in gww
pico_mid <- (4/3)*pi*((0.5*0.0001*5.1)^3)
large_mid <- (4/3)*pi*((0.5*0.0001*105)^3)
micro_mid <- (4/3)*pi*((0.5*0.0001*101)^3)
meso_mid <- (4/3)*pi*((0.5*0.0001*10100)^3)
```

## Convert to gww and then abundance

```{r}
# Get numerical abundance by dividing by size class mid point
# This step assumes that all plankton are the midpoint size
pico_abund <- df_pico_long[,2]/pico_mid
large_abund <- (df_diat_long[,2] + df_diaz_long[,2])/large_mid
micro_abund <- df_zmicro_long[,2]/micro_mid
meso_abund <- df_zmeso_long[,2]/meso_mid

# Combine mid-point sizes for generating the x-axis for the linear fit
plankton_x <- log10(c(pico_mid, micro_mid, large_mid, meso_mid))

# The full spectrum sizes were generated by setting up a mizer params:
```

## Load mizer

```{r}
library(mizer)
# params <- newMultispeciesParams(mizer::NS_params@species_params, min_w_pp = 1e-14)
```

## Create model params

We need this to get the full size spectrum from the model

```{r}
# The full spectrum sizes were generated by setting up a mizer params:
# HIparams <- read.csv("HIregion_species_params.csv")
# HIinter <- read.csv("HIregion_inter.csv")[,-1]
# rownames(HIinter) <- colnames(HIinter)
# 
# params <- newMultispeciesParams(HIparams, interaction = HIinter, kappa = 1e12, min_w_pp = 1e-14)

params17 <- readRDS("~/Desktop/2023 SO MODEL PAPER/HIMI/params/params17.rds")
HIMIparams <- params17@params
HIMIinter <- params17@params@interaction

# and accessing the full size range
full_x <- log10(HIMIparams@w_full)
length(full_x)

#Does this mean I have 209 size classes?
```


## Create background resource

```{r}
# Creating background resource for full_x, using the actual slope and intercept from the linear models.
# Create array and fill it
# out_isimip <- array(numeric(), c(600,226)) # 600 time steps by 226 size classes
# isimip_slope <- array(numeric(), c(600,1)) # 600 time steps
# isimip_intercept <- array(numeric(), c(600,1)) # 600 time steps

# Create array and fill it
out_isimip <- array(numeric(), c(600,209)) # 600 time steps by 209 size classes
isimip_slope <- array(numeric(), c(600,1)) # 600 time steps
isimip_intercept <- array(numeric(), c(600,1)) # 600 time steps

# y values
for (t in seq(1,600,1)) {
	isimip_plankton <- log10(c(pico_abund$total_C_g[t], micro_abund$total_C_g[t], large_abund$total_C_g[t], meso_abund$total_C_g[t]))
		
	# Calculate slope and intercept, expand spectra for full size range
	# Linear fits
	isimip_lm <- lm(isimip_plankton ~ plankton_x)
	
	# Expand to full size range
	# out_isimip[t,] <- isimip_lm$coefficients[2] * full_x + isimip_lm$coefficients[1]
	out_isimip[t,] <- isimip_lm$coefficients[2]*1.03 * full_x + isimip_lm$coefficients[1]*0.85
	# The scaling for the slope and intercept were determined following the method in 
	# Woodworth-Jefcoats et al. (2019)  More information is provided below.
	
	# Save slope and intercept, for diagnostics
	isimip_intercept[t,1] <- isimip_lm$coefficients[1]
	isimip_slope[t,1] <- isimip_lm$coefficients[2]
	
}
```

## Plot full spectrum

```{r}
ggplot(,aes(x = full_x, y = out_isimip[1,])) +
  geom_point() +
  # geom_line() +
  # geom_smooth() +
  scale_y_log10() +
  xlab("Size (log10 g)") +
  ylab("Abundance (log10)") +
  theme_bw()
```

## Intercept timeseries 

```{r}
months <- seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months") # create a tidy date variable

ggplot(,aes(x = months, y = isimip_intercept[,1])) +
  geom_point() +
  geom_line() +
  geom_smooth() +
  xlab("Year") +
  ylab("log10 Intercept") +
  theme_bw()
```

## Slope timeseries

```{r}
ggplot(,aes(x = months, y = isimip_slope[,1])) +
  geom_point() +
  geom_line() +
  geom_smooth() +
  xlab("Year") +
  ylab("Background Resource Slope") +
  theme_bw()
```

## Save files

```{r}
# Save
# write.table(out_isimip, file = "GFDL_resource_spectra_60arcmin.dat", quote = FALSE, row.names = TRUE, col.names = TRUE)
# write.table(isimip_slope, file = "GFDL_resource_slope_60arcmin.dat", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
# write.table(isimip_intercept, file = "GFDL_resource_intercept_60arcmin.dat", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)

write.table(out_isimip, file = "GFDL_resource_spectra_15arcmin.dat", quote = FALSE, row.names = TRUE, col.names = TRUE)
write.table(isimip_slope, file = "GFDL_resource_slope_15arcmin.dat", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
write.table(isimip_intercept, file = "GFDL_resource_intercept_15arcmin.dat", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)


```

#AE: need to check all of above, I just did a quick run through and seemed to all be doing its thing. Then temperature preparation next!
