---
title: toothfish project
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(mizerHowTo)
library(therMizer)
library(parallel)
library(tictoc)
load("inputs/size_spec_inputs.RData")
```


# Helpful functions

```{r}
# Calculate the temperature scaling factor for the encounter rate and 
# predation rate
scaled_temp_fun <- function(params, t) {
    # Using t+1 to avoid calling ocean_temp[0,] at the first time step

  # print("model time")
  # print(t)
  # print("adjusted time")
  # print(t + params@other_params$other$t_idx)
  # print("temperature")
  # print(other_params(params)$ocean_temp)
    if(t == 0) temp_at_t <- other_params(params)$ocean_temp[t + 1,]
    else temp_at_t <- other_params(params)$ocean_temp[t + params@other_params$other$t_idx,]

    # print("temperature in scaled_temp")
    # print(temp_at_t)
    
    # Calculate unscaled temperature effect using a generic polynomial rate equation
    unscaled_temp_effect <- 
        temp_at_t * (temp_at_t - species_params(params)$temp_min) * 
        (species_params(params)$temp_max - temp_at_t)
    
    # Scale using new parameter
    scaled_temp_effect <- 
        unscaled_temp_effect / species_params(params)$encounterpred_scale
    # print("intermediate temp effect")
    # print(scaled_temp_effect)
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    scaled_temp_effect[above_max | below_min] = 0
    
    # print("temp effect")
    # print(scaled_temp_effect)
    
    return(scaled_temp_effect)
}


therMizerEncounter <- function(params, t, ...) {
    
      # Calculate maximum possible encounter rate
      max_encounter <- mizerEncounter(params, t, ...)
      
      # Apply temperature effect
      return(max_encounter * scaled_temp_fun(params, t))
      
}

therMizerPredRate <- function(params, t, ...) {
      # Calculate maximum possible encounter rate
      max_predrate <- mizerPredRate(params, t, ...)
      # print("predrate")
      # print(max_predrate)
      # Apply temperature effect
      return(max_predrate * scaled_temp_fun(params, t))
      
}


therMizerEReproAndGrowth <- function(params, t, encounter, feeding_level, ...) {
    
  # print("t")
  # print(t)
  # print("emcounter")
  # print(encounter)
  # print("feeding")
  # print(feeding_level)
  
    # Using t+1 to avoid calling ocean_temp[0,] at the first time step
    temp_at_t <- other_params(params)$ocean_temp[t + params@other_params$other$t_idx,]
  
    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))
    
    # Arrhenius equation scaled to a value between 0 and 1
    temp_effect_metabolism <- 
        (unscaled_temp_effect - species_params(params)$metab_min) /
        species_params(params)$metab_range
    
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    temp_effect_metabolism[above_max | below_min] = 0
  
  # Apply scaled Arrhenius value to metabolism
    sweep((1 - feeding_level) * encounter, 1,
          species_params(params)$alpha, "*", check.margin = FALSE) - 
        metab(params)*temp_effect_metabolism  
      
}

# Set up new resource forcing "function" - just changes to different time slot in the n_pp_array array
plankton_forcing <- function(params, t, ...) {
  w_cut_off <- 10
  pkt <- 10^(other_params(params)$n_pp_array[t + params@other_params$other$t_idx,])/params@dw_full # converting to density
  pkt[which(as.numeric(names(pkt)) >= w_cut_off)] <- 0
  
  return(pkt)  
}

plotDiet2 <- function (sim, species = NULL, xlim = c(1, NA), returnData = F) 
{
  params <- sim@params
  diet <- getDiet(params)
  plot_dat <- melt(diet)
  plot_dat <- plot_dat[plot_dat$value > 0, ]
  colnames(plot_dat) <- c("Predator", "size", "Prey", 
                          "Proportion")
  legend_levels <- intersect(names(params@linecolour), unique(plot_dat$Prey))
  if (is.null(species)) 
    p <- ggplot(plot_dat) + facet_wrap(. ~ Predator, scales = "free")
  else p <- ggplot(filter(plot_dat, Predator == species))
  p <- p + 
    geom_area(aes(x = size, y = Proportion, fill = Prey)) + 
    scale_x_continuous(limits = c(1, NA), name = "Size [g]", trans = "log10") + 
    scale_fill_manual(values = sim@params@linecolour[legend_levels]) + 
    theme(legend.position = "right", 
          legend.key = element_rect(fill = "white"), 
          panel.background = element_blank(),
          panel.grid.minor = element_line(color = "gray"), 
          strip.background = element_blank())
  if (returnData) 
    return(plot_dat)
  else return(p)
}
```

# equation for fisheries yield

$$yield = catchability\cdot selectivity \cdot effort \cdot biomass   $$


# Setting up params objects

The ecosystem is calibrated and the time series to force planktong and temperature are ready. Time to combine everything.

Loading params without any temperature so easy to edit with block below

```{r}
sim <- readRDS("params/params10.RDS")
params <- sim@params
first_year <- 1950
last_year <- 2093
times <- seq(first_year,last_year, by = 1)

## Setting up temperature range

species_params(params)$temp_min <- c(1.8,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.8,1.6) # fishbase
# species_params(params)$temp_min <- c(1.4,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.4,1.4) # adjusted so species survive historical temperatures

# species_params(params)$temp_max <- c(8.8,2.9,2.5,2.8,1.4,2,2.3,3.1,3.9) #fishbase
species_params(params)$temp_max <-   c(8.8,2.9,2.5,2.8,2.4,2.4,2.3,3.1,3.9) # adjusted


## Time index
other_params(params)$t_idx = -1949 # temperature array starts at 1950 so any time generated by the effort array will match the rownumber of the temperature array

## Create parameter for scaling encounter and mortality rates so scalar is always between 0 and 1
species_params(params)$encounterpred_scale <- 
    rep(NA, nrow(species_params(params)))

for (indv in seq(1:length(species_params(params)$temp_min))) {
    
    # Create a vector of all temperatures each species might encounter
    temperature <- seq(species_params(params)$temp_min[indv], 
                       species_params(params)$temp_max[indv], 
                       by = 0.1)
    
    # Find the maximum value of the unscaled effect of temperature on encounter
    # and predation rate for each species 
  species_params(params)$encounterpred_scale[indv] <- 
      max((temperature) * 
              (temperature - species_params(params)$temp_min[indv]) *
              (species_params(params)$temp_max[indv] - temperature)
          )
}

## Determine the minimum, maximum, and range of value for the effect of 
# temperature on metabolism
    
min_metab_value <- 
    (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_min)))))
max_metab_value <- 
    (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_max)))))
    
species_params(params)$metab_min <- min_metab_value
species_params(params)$metab_range <- max_metab_value - min_metab_value

```


Set the new rate functions and new resource

```{r}

params <- setRateFunction(params, "Encounter", "therMizerEncounter")
params <- setRateFunction(params, "PredRate", "therMizerPredRate")
params <- setRateFunction(params, "EReproAndGrowth", "therMizerEReproAndGrowth")

params <- setResource(params, resource_dynamics = "plankton_forcing")

```
Adding fisheries to the mix

```{r}

ll_area_yr <- read_csv("inputs/ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)
ll_km2$catchT <- ll_km2$catchT * 1e6 #in grams
ll_avg <- mean(ll_km2$catchT)
effort_avg <-mean(ll_km2$effort)

effort <-array(ll_km2$effort, dim = c(length(ll_km2$effort),1), 
                 dimnames = list("time" = ll_km2$year, "gear" = params@gear_params$gear[1]))

```


divide ll_effort and catch by total area to get effort and catch per km2. the total area is from the mask in FISHMIP folder


Set the scenarios

```{r}

# Create parameter objects
params_IPSL_picontrol <- params
params_IPSL_ssp1rcp26 <- params
params_IPSL_ssp5rcp85 <- params

# Attach temperature
other_params(params_IPSL_picontrol)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_PIcontrol.RDS")
other_params(params_IPSL_ssp1rcp26)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_126.RDS")
other_params(params_IPSL_ssp5rcp85)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_585.RDS")

# Attach plankton
other_params(params_IPSL_picontrol)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_PIcontrol.RDS")
other_params(params_IPSL_ssp1rcp26)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_126.RDS")
other_params(params_IPSL_ssp5rcp85)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_585.RDS")


```




```{r}
sim <- project(params_IPSL_ssp1rcp26, effort = effort)
```

simulation using historical data works well, picontrol doesn't


```{r}

mizer::plotBiomass(sim)
mizer::plotSpectra(sim)

```


```{r}
res <- mizer::plotYield(sim,return_data = T)
res$ObservedYield <- ll_km2$catchT

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")
```


# New calibration

Redo previous calibration with temperature 


```{r}
getErrorCustom <- function(vary, params, dat, times, tol = 0.001, 
    timetorun = 10)
{
  # params@species_params$R_max[1:9]<-10^vary[1:9]
  # params@species_params$erepro[1:9]<-vary[10:18]
  # params@species_params$interaction_resource[1:9] <- vary[19:27]
  # 
  # params <- setParams(params)
  # 
  # interaction <- params@interaction
  # interaction[] <- matrix(vary[28:108],nrow = 9) # stop at 54 if looking only at 3 biggest species
  # 
  # params <- setInteraction(params,interaction)
  
  params@species_params$erepro[1] <- vary

    # params <- projectToSteady(params, distance_func = distanceSSLogN, 
    #     tol = tol, t_max = length(times), return_sim = F)
    
    sim <- project(params,t_max = length(times),t_start = 1950, progress_bar = F,initial_n =  params@initial_n, initial_n_pp = params@other_params$other$n_pp_array[1,])
    
 # get biomass through time
  biomass <- sweep(sim@n, 3, sim@params@w * sim@params@dw, "*")
  biomass <- biomass[which(dimnames(biomass)$time == "2003"):which(dimnames(biomass)$time == "2020"),,] # trageting fishing period
  
  #get yield through time from model:
  
  f_gear<-mizer::getFMortGear(params,effort)
  # f_gear <- f_gear[1:13,,,,drop=F]
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred[1:13]) - log(obs[1:13]))^2,na.rm=T)
  
plot_dat <- data.frame(obs,pred)
plot_dat$Year <- 2003:2020

p <- ggplot(plot_dat, aes(x = Year)) +
  geom_line(aes(y = pred)) +
  geom_line(aes(y = obs), color = "red")
print(p)

p2 <- mizer::plotBiomass(sim)
print(p2)
    return(error)
}



```

```{r}
tic()
dat <- ll_km2

  effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

# create set of params for the optimisation process
params_optim <- sim@params
# params_optim@other_params$other$t_idx <- 1  # t_Start doesn't work in project to steady so always starts at 0
# vary<-c(log10(params_optim@species_params$R_max),
        # params_optim@species_params$erepro,
        # params_optim@gear_params$catchability[1])


# vary<-c(log10(params_optim@species_params$R_max),
#         params_optim@species_params$erepro,
#         params_optim@species_params$interaction_resource,
#         params_optim@interaction)

vary <- params_optim@species_params$erepro


params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim, dat = ll_km2, effort = effort, times = times,
                                             method   ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-10,9),rep(.1,9),rep(.1,81)),
                                             upper= c(rep(15,9),rep(1,9),rep(.99,9),rep(.99,81)),
                                             parallel=list(loginfo=TRUE, forward=TRUE),
                                             plankton_forcing, therMizerEncounter, therMizerPredRate, therMizerEReproAndGrowth, scaled_temp_fun)
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_temperature.RDS")
```

Not working, doing it manually


```{r}
params_optim <- sim@params

tempParams <- params_optim

tempParams@species_params$erepro[1] <- vary <- params_optim@species_params$erepro[1]*10
sim <- project(tempParams,t_max = length(times),t_start = 1950, progress_bar = F,initial_n =  tempParams@initial_n, initial_n_pp = tempParams@other_params$other$n_pp_array[1,])
mizer::plotBiomass(sim)


res <- getErrorCustom(params = params_optim,vary = vary,dat = ll_km2,times = times)

res
```

Can't get yield to match post 2015
Decreasing erepro scaled down the entire curve but keep same profile
Increasing erepro doesn't change anything

The drop is due to something else. Preys?


# Predictions

Constant effort


```{r}

effort_cst <- array(data = c(effort,rep(effort[13],(2093-2002-18))), dim = c(2093-2002,1), dimnames = list("time" = 2003:2093, "gear" = "longline"))

sim <- project(params_IPSL_ssp1rcp26, effort = effort_cst)

mizer::plotBiomass(sim)
mizer::plotSpectra(sim)

res <- mizer::plotYield(sim,return_data = T)
res$ObservedYield <- c(ll_km2$catchT, rep(NA,(2093-2002-18)))

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")

```

No fishing

```{r}


sim <- project(params_IPSL_ssp1rcp26, 
               initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      t_max = length(times),t_start = 1950)

mizer::plotBiomass(sim)
mizer::plotSpectra(sim, time_range = 2045)


```

```{r}


sim <- project(params_IPSL_ssp5rcp85, 
               initial_n_pp = params_IPSL_ssp5rcp85@other_params$other$n_pp_array[1,],
                                      t_max = length(times),t_start = 1950)

mizer::plotBiomass(sim)
mizer::plotSpectra(sim)
plotDiet2(sim)

```


# Plots

6.	%Change in biomass for all species under projections relative to biomass averaged over recent period (2000s?).
7.	%Change in catches for toothfish under projections relative to biomass averaged over recent period (2000s?).
8.	%Change in mean body size for all species under projections relative to biomass averaged over recent period (2000s?).


```{r}
myTheme <- function(){
     theme(
        legend.key = element_rect(fill = "white"),
        legend.position = "bottom",
        panel.background = element_rect(fill = "white", color = "black"),
        strip.background = element_blank(),
        panel.grid.minor = element_line(colour = "grey92"))
}

fig_width = 15
fig_height =10
```



### ssp126

calculating average biomass for each species from 2000 to 2009

```{r}

sim <- project(params_IPSL_ssp1rcp26,initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      t_max = 10,t_start = 2000)

biomass <- mizer::plotBiomass(sim, return_data = T)
biom_avg <- NULL
for(iSpecies in sim@params@species_params$species){
  temp <- filter(biomass, Species == iSpecies)
  
  biom_avg <- rbind(biom_avg, mean(temp$Biomass))
}

biom_avg <- data.frame(Species = sim@params@species_params$species, biomass = biom_avg)

```

Projecting scenario in time

```{r}


sim <- project(params_IPSL_ssp1rcp26,initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      t_max = length(times),t_start = 1950)

biomass <- mizer::getBiomass(sim)

biomass_relative <- apply(biomass,1,FUN = function(x){x/biom_avg$biomass})

plot_dat <- reshape2::melt(biomass_relative)
colnames(plot_dat)[1] <- "Species"


p1 <- ggplot(plot_dat)+
  geom_line(aes(x = time, y = value, color = Species, size = Species)) +
  scale_y_continuous("Proportional change in biomass") +
  scale_x_continuous("Time in years") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_color_manual(values = params_IPSL_ssp5rcp85@linecolour[1:9]) +
  scale_size_manual(values = c(2,rep(.8,8)))+
  myTheme()
p1

ggsave(p1, file = "figures/biomass26.png", units = "cm", width = fig_width, height = fig_height)
```


Yield when assuming constant effort (2015) onwards

```{r}
effort_cst <- array(data = c(effort,rep(effort[13],(2093-2002-18))), dim = c(2093-2002,1), dimnames = list("time" = 2003:2093, "gear" = "longline"))

sim <- project(params_IPSL_ssp1rcp26, effort = effort_cst)

mizer::plotBiomass(sim)
mizer::plotSpectra(sim)

res <- mizer::plotYield(sim,return_data = T)
res$ObservedYield <- c(ll_km2$catchT, rep(NA,(2093-2002-18)))

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")
```


Yield relative change to average 2000-2009

```{r}

sim <- project(params_IPSL_ssp1rcp26,initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      t_max = 91,t_start = 2003)

yield <- mizer::getYield(sim)
yield_avg <- mean(yield[,1])


```


Projecting in time

```{r}

sim <- project(params_IPSL_ssp1rcp26,initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      effort = effort_cst)

yield <- mizer::getYield(sim)

yield_relative <-yield[,1]/yield_avg

plot_dat <- data.frame(Year = as.numeric(names(yield_relative)), yield = yield_relative)


p2 <- ggplot(plot_dat)+
  geom_line(aes(x = Year, y = yield)) +
  geom_hline(yintercept = 1, linetype = "dashed")
p2


```


mean body size - 2003-2012 (including fisheries)
using getMeanWeigth that does sum biomass / sum numbers

```{r}

mean_size <- NULL
for(iSpecies in sim@params@species_params$species){
  
  temp <- mizer::getMeanWeight(sim,species = iSpecies)
  mean_size <- cbind(mean_size,temp)
}

colnames(mean_size) <- sim@params@species_params$species

mean_size_avg <- apply(mean_size[1:10,],2,mean)

relative_size <- sweep(mean_size,2,mean_size_avg,"/")

plot_dat <- reshape2::melt(relative_size)
colnames(plot_dat) <- c("Year", "Species", "mean_size")


p3 <- ggplot(plot_dat)+
  geom_line(aes(x = Year, y = mean_size, color = Species, size = Species)) +
    scale_y_continuous(name = "Proportional change in mean body size") +
  scale_x_continuous("Time in years") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_color_manual(values = params_IPSL_ssp5rcp85@linecolour[1:9]) +
  scale_size_manual(values = c(2,rep(.8,8)))+
  myTheme()
p3

ggsave(p3, file = "figures/meanSize26.png", units = "cm", width = fig_width, height = fig_height)

```

### ssp585

calculating average biomass for each species from 2000 to 2009

```{r}

sim <- project(params_IPSL_ssp5rcp85,initial_n_pp = params_IPSL_ssp5rcp85@other_params$other$n_pp_array[1,],
                                      t_max = 10,t_start = 2000)

biomass <- mizer::plotBiomass(sim, return_data = T)
biom_avg <- NULL
for(iSpecies in sim@params@species_params$species){
  temp <- filter(biomass, Species == iSpecies)
  
  biom_avg <- rbind(biom_avg, mean(temp$Biomass))
}

biom_avg <- data.frame(Species = sim@params@species_params$species, biomass = biom_avg)

```

Projecting scenario in time

```{r}


sim <- project(params_IPSL_ssp5rcp85,initial_n_pp = params_IPSL_ssp5rcp85@other_params$other$n_pp_array[1,],
                                      t_max = length(times),t_start = 1950)

biomass <- mizer::getBiomass(sim)

biomass_relative <- apply(biomass,1,FUN = function(x){x/biom_avg$biomass})

plot_dat <- reshape2::melt(biomass_relative)
colnames(plot_dat)[1] <- "Species"


p4 <- ggplot(plot_dat)+
  geom_line(aes(x = time, y = value, color = Species, size = Species)) +
  scale_y_continuous("Proportional change in biomass") +
  scale_x_continuous("Time in years") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_color_manual(values = params_IPSL_ssp5rcp85@linecolour[1:9]) +
  scale_size_manual(values = c(2,rep(.8,8)))+
  myTheme()
p4

ggsave(p4, file = "figures/biomass85.png", units = "cm", width = fig_width, height = fig_height)

```


Yield when assuming constant effort (2015) onwards

```{r}
effort_cst <- array(data = c(effort,rep(effort[13],(2093-2002-18))), dim = c(2093-2002,1), dimnames = list("time" = 2003:2093, "gear" = "longline"))

sim <- project(params_IPSL_ssp5rcp85, effort = effort_cst)

mizer::plotBiomass(sim)
mizer::plotSpectra(sim)

res <- mizer::plotYield(sim,return_data = T)
res$ObservedYield <- c(ll_km2$catchT, rep(NA,(2093-2002-18)))

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")
```


Yield relative change to average 2000-2009

```{r}

sim <- project(params_IPSL_ssp5rcp85,initial_n_pp = params_IPSL_ssp5rcp85@other_params$other$n_pp_array[1,],
                                      t_max = 91,t_start = 2003)

yield <- mizer::getYield(sim)
yield_avg <- mean(yield[,1])


```


Projecting in time

```{r}

sim <- project(params_IPSL_ssp5rcp85,initial_n_pp = params_IPSL_ssp5rcp85@other_params$other$n_pp_array[1,],
                                      effort = effort_cst)

yield <- mizer::getYield(sim)

yield_relative <-yield[,1]/yield_avg

plot_dat <- data.frame(Year = as.numeric(names(yield_relative)), yield = yield_relative)


p5 <- ggplot(plot_dat)+
  geom_line(aes(x = Year, y = yield)) +
  geom_hline(yintercept = 1, linetype = "dashed")

p5

```


mean body size - 2003-2012 (including fisheries)
using getMeanWeigth that does sum biomass / sum numbers

```{r}

mean_size <- NULL
for(iSpecies in sim@params@species_params$species){
  
  temp <- mizer::getMeanWeight(sim,species = iSpecies)
  mean_size <- cbind(mean_size,temp)
}

colnames(mean_size) <- sim@params@species_params$species

mean_size_avg <- apply(mean_size[1:10,],2,mean)

relative_size <- sweep(mean_size,2,mean_size_avg,"/")

plot_dat <- reshape2::melt(relative_size)
colnames(plot_dat) <- c("Year", "Species", "mean_size")


p6 <- ggplot(plot_dat)+
  geom_line(aes(x = Year, y = mean_size, color = Species, size = Species)) +
    scale_y_continuous(name = "Proportional change in mean body size") +
  scale_x_continuous("Time in years") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_color_manual(values = params_IPSL_ssp5rcp85@linecolour[1:9]) +
  scale_size_manual(values = c(2,rep(.8,8)))+
  myTheme()

p6

ggsave(p6, file = "figures/meanSize85.png", units = "cm", width = fig_width, height = fig_height)

```



facet of temperature profile per species
add dots for average historical data (last 20 years) and average increase (last 20 years)




```{r}
# temperature profile

temp_vec <- seq(-1.5,9,.05)
scalar <- NULL
scalarM <- NULL

for(iTemp in temp_vec){
  
temp_at_t <- iTemp

# temp_at_t <- temp_at_t + 273

# unscaled_temp_effect <- 
#   temp_at_t * (temp_at_t - (species_params(params)$temp_min + 273)) * 
#   (species_params(params)$temp_max - temp_at_t + 273)

unscaled_temp_effect <- 
  temp_at_t * (temp_at_t - (species_params(params)$temp_min)) * 
  (species_params(params)$temp_max - temp_at_t)

scaled_temp_effect <- 
  unscaled_temp_effect / species_params(params)$encounterpred_scale

above_max <- temp_at_t > species_params(params)$temp_max
below_min <- temp_at_t < species_params(params)$temp_min
scaled_temp_effect[above_max | below_min] = 0
# scaled_temp_effect[scaled_temp_effect < 0] = 0

scalar <- rbind(scalar,scaled_temp_effect)


    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))
    
    # Arrhenius equation scaled to a value between 0 and 1
    temp_effect_metabolism <- 
        (unscaled_temp_effect - species_params(params)$metab_min) /
        species_params(params)$metab_range
    
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    temp_effect_metabolism[above_max | below_min] = 0
    
    scalarM <- rbind(scalarM, temp_effect_metabolism)

}

rownames(scalar) <- temp_vec
colnames(scalar) <- params@species_params$species
plot_dat <- reshape2::melt(scalar)
colnames(plot_dat) <- c("temperature","Species","scalar")
plot_dat$Type <- "Encounter"

# removing scalar = 0 except at the beginning and end of curve
zero_keep <- NULL
zero_val <- which(plot_dat$scalar == 0)
for(i in 1:(length(zero_val)-1)){
  if(zero_val[i+1] != zero_val[i] + 1){
    zero_keep <- c(zero_keep, i, i+1)
  }
}
# zero keep has the position of the row to keep in zero_val
zero_val <- zero_val[-zero_keep]
plot_dat <-plot_dat[- zero_val,] 

rownames(scalarM) <- temp_vec
colnames(scalarM) <- params@species_params$species
plot_datM <- reshape2::melt(scalarM)
colnames(plot_datM) <- c("temperature","Species","scalar")
# plot_datM$scalar[plot_datM$scalar == 0] <- NA
plot_datM$Type <- "Metabolism"

# removing scalar = 0 except at the beginning of curve
zero_keep <- NULL
zero_val <- which(plot_datM$scalar == 0)
for(i in 1:(length(zero_val)-1)){
  if(zero_val[i+1] != zero_val[i] + 1){
    zero_keep <- c(zero_keep, i)
  }
}
# zero keep has the position of the row to keep in zero_val
zero_val <- zero_val[-zero_keep]
plot_datM <-plot_datM[- zero_val,] 

plot_dat <- rbind(plot_dat, plot_datM)

# plot_dat[which(plot_dat$scalar == 0),]

plot_dat <- plot_dat[-780,] # for some reason that one value really wants to stay in the df

# points for current and projected temperature
temperature <- params_IPSL_ssp1rcp26@other_params$other$ocean_temp[,1]
historical <- mean(temperature[40:60])
projection <- mean(temperature[124:144])
scalar <- NULL
scalarM <- NULL
for(iTemp in c(historical,projection)){
  
temp_at_t <- iTemp

# temp_at_t <- temp_at_t + 273

# unscaled_temp_effect <- 
#   temp_at_t * (temp_at_t - (species_params(params)$temp_min + 273)) * 
#   (species_params(params)$temp_max - temp_at_t + 273)

unscaled_temp_effect <- 
  temp_at_t * (temp_at_t - (species_params(params)$temp_min)) * 
  (species_params(params)$temp_max - temp_at_t)

scaled_temp_effect <- 
  unscaled_temp_effect / species_params(params)$encounterpred_scale

above_max <- temp_at_t > species_params(params)$temp_max
below_min <- temp_at_t < species_params(params)$temp_min
scaled_temp_effect[above_max | below_min] = 0
scaled_temp_effect[scaled_temp_effect < 0] = 0

scalar <- rbind(scalar,scaled_temp_effect)


    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))
    
    # Arrhenius equation scaled to a value between 0 and 1
    temp_effect_metabolism <- 
        (unscaled_temp_effect - species_params(params)$metab_min) /
        species_params(params)$metab_range
    
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    temp_effect_metabolism[above_max | below_min] = 0
    
    scalarM <- rbind(scalarM, temp_effect_metabolism)

}

rownames(scalar) <- c(historical,projection)
colnames(scalar) <- params@species_params$species
plot_point <- reshape2::melt(scalar)
plot_point$Scenario <- c("Historical","Projection")
colnames(plot_point)[1:3] <- c("temperature","Species","scalar")
plot_point$Type <- "Encounter"

rownames(scalarM) <- c(historical,projection)
colnames(scalarM) <- params@species_params$species
plot_pointM <- reshape2::melt(scalarM)
plot_pointM$Scenario <- c("Historical","Projection")
colnames(plot_pointM)[1:3] <- c("temperature","Species","scalar")
plot_pointM$Type <- "Metabolism"

plot_point <- rbind(plot_point, plot_pointM)

p7 <- ggplot(plot_dat)+
  geom_line(aes(x = temperature, y = scalar, color = Type)) +
  geom_point(data = plot_point, aes(x = temperature, y = scalar, shape = Scenario, color = Scenario), size = 2) +
  scale_x_continuous("Temperature in C")+
  scale_y_continuous("Scalar value") +
  facet_wrap(~Species, scales = "free") +
myTheme() +
  guides(color=guide_legend(nrow=2,byrow=TRUE), shape = guide_legend(nrow=2,byrow=TRUE)) +
  scale_color_manual(values = c("dodgerblue","darkgreen","green3","darkorange2"))
  
p7

ggsave(p7, file = "figures/tempProfile.png", units = "cm", width = fig_width, height = fig_height)


```

Catch predictions of both scenarios

```{r}
effort_cst <- array(data = c(effort,rep(effort[13],(2093-2002-18))), dim = c(2093-2002,1), dimnames = list("time" = 2003:2093, "gear" = "longline"))

sim85 <- project(params_IPSL_ssp5rcp85, effort = effort_cst)
res <- mizer::plotYield(sim85,return_data = T)
res$Scenario <- "SSP8.5"
sim26 <- project(params_IPSL_ssp1rcp26, effort = effort_cst)
res26 <- mizer::plotYield(sim26,return_data = T)
res26$Scenario <- "SSP2.6"

ObservedYield <- data.frame(Year =  2003:2093, Yield =  c(ll_km2$catchT, rep(NA,(2093-2002-18))), Species = "D.ele")
ObservedYield$Scenario <- "Observed yield"

plot_dat <- rbind(res,res26,ObservedYield)
plot_dat$Yield <- plot_dat$Yield/1e6 # conversion to tons

p8 <- ggplot(plot_dat) +
  geom_line(aes(x = Year, y = Yield, color = Scenario)) +
  scale_x_continuous("Time in years") +
  scale_y_continuous("Yield in tonnes") +
  myTheme()
  
p8

ggsave(p8, file = "figures/futureCatch.png", units = "cm", width = fig_width, height = fig_height)

```



Proportional change in n_pp for both scenarios


```{r}


hist_avg <- apply(sim26@params@other_params$other$n_pp_array[40:60,],2,mean)
rcp26_avg <- apply(sim26@params@other_params$other$n_pp_array[124:144,],2,mean)
rcp85_avg <- apply(sim85@params@other_params$other$n_pp_array[124:144,],2,mean)

npp26 <- rcp26_avg/hist_avg
npp85 <- rcp85_avg/hist_avg

plot_dat <- data.frame(value = c(npp26,npp85), Scenario = c(rep("SSP2.6",209),rep("SSP8.5",209)), Size = rep(as.numeric(names(npp26)),2))

p9 <- ggplot(plot_dat) +
  geom_line(aes(x = Size, y = value, color = Scenario)) +
  scale_x_continuous(name = "Size in [g]", trans = "log10") +
  scale_y_continuous("Proportional change of biomass") +
  myTheme()
p9

ggsave(p9, file = "figures/nppChange.png", units = "cm", width = fig_width, height = fig_height)

```

historical diets - pre temperature forcing

```{r}
sim <- readRDS("params/params10.RDS")

p10 <- plotDiet2(sim, species = "D.ele")

p10

ggsave(p10, file = "figures/diet.png", units = "cm", width = fig_width, height = fig_height)

```



historical growth - pre temperature forcing

```{r}

sim <- readRDS("params/params6.RDS")

p11 <- plotGrowthCurves(sim, species_panel = T)

p11

ggsave(p11, file = "figures/growth.png", units = "cm", width = fig_width, height = fig_height)

```


Preping the panel plots

```{r}
library(ggpubr)
```


All custom functions are in report2.Rmd

Figure 1 size spectrum and biomass calibration
```{r}
# using sim before fishing time series calibration to compare biomass to data using only average fishing

sim21 <- readRDS("params/params9.RDS")

params <- sim21@params

ll_area_yr <- read_csv("ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)
ll_km2$catchT <- ll_km2$catchT * 1e6 #in grams
ll_avg <- mean(ll_km2$catchT)
effort_avg <-mean(ll_km2$effort)

  effort <-array(ll_km2$effort, dim = c(length(ll_km2$effort),1), 
                 dimnames = list("time" = ll_km2$year, "gear" = params@gear_params$gear[1]))

params@species_params$yield_observed <- c(ll_avg,rep(0,8))

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(7e-7,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll
initial_effort(params) <- effort_avg # somehow doesn't work with gear_params

sim22 <- project(params, t_max = 200)

plotdat12 <- plotBiomassObservedVsModelCustom(sim22)

plotdat13 <- mizer::plotSpectra(sim22)

p12 <- plotdat12 + myTheme()

p13 <- plotdat13 + myTheme()

p14 <- ggarrange(p13, p12, ncol=2, common.legend = TRUE, legend="bottom", labels = c('A', 'B'))

ggsave(p14, filename = "figures/figure1.png", units = "cm", width = 20, height = 15)

```


Figure 2 Growth and diet

```{r}

p10 <- p10 + myTheme()
p11 <- p11 + myTheme()

p15 <- ggarrange(p10, p11, ncol=2, legend="bottom", labels = c('A', 'B'))
ggsave(p15, filename = "figures/figure2.png", units = "cm", width = 20, height = 15)


```

Figure 6 pannel of relative change

```{r}

p4 <- p4 + scale_y_continuous(name = "")
p6 <- p6 + scale_y_continuous(name = "")

p16 <- ggarrange(p1, p4, p3, p6, ncol=2, nrow = 2, common.legend = TRUE, legend="bottom", labels = c('A', 'B','C','D'))

ggsave(p16, filename = "figures/figure3.png", units = "cm", width = 20, height = 20)

```

