---
title: doing the same calibration but better
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(terra)
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(therMizer)
library(parallel)
library(tictoc)
source("functions.R")

load("inputs/size_spec_inputs.RData")
```


# The data

I have two datasets. The biomass survey is RSTS data which is kg of species per area surveyed (effort column in km2). The longlines concerns the toothfish only and is in number of hooks (effort) and catches in tonnes. The data for the longline is for the whole area whereas the survey is per km2. The first step is to get everything to the same scale, in our case, the whole studied area.

# The studied area

Nicole sent an HIMI mask that I can use to get the total surface area

```{r}
HIMI_shape<-sf::st_read("HIMI_Mask/HIMI_EEZ_polygon/HIMI_EEZ_polygon.shp") # the shape file of the area
plot(HIMI_shape)
HIMI_area <- sf::st_area(HIMI_shape)
```

The surface area in km2 is

```{r}

total_area <- as.numeric(HIMI_area)/1e6
total_area

```

Now I need the fraction of surveyed area compared to the total area. This will allow me to model the shallows area with the juveniles within the bigger area.

The details of the area surveyed is in Area_surveyed.Rdata

```{r}

load("inputs/Area_surveyed.RData")

```


Now I can estimate the biomass for each species each year in the whole shallows. Hopefully, the sizes not caught by the survey but modelled by Mizer will match well what happens outside of the shallows


```{r}

rsts_total <- rsts_trends[,-(1:2)] * year_tot[3:16,]$tot_area / rsts_trends$Effort

rsts_total$year <- rsts_trends$Year

saveRDS(rsts_total, file = "inputs/rsts_total.rds")
```


Need to re-calibrate the model with these values. Are they more or less constant?

```{r}

plot_dat <- reshape2::melt(rsts_total)
plot_dat$year <- as.numeric(plot_dat$year)

ggplot(plot_dat) +
  geom_line(aes(x = year, y = value, color = variable)) +
  facet_wrap(~variable, scales = "free") +
  scale_y_continuous(trans = "log10")

```

Big no. Let's average the values for the calibration and plot standard deviation on the final plot

# Biomass calibration

The latest saved param object is params10


```{r}

sim1 <- readRDS("params/params10.RDS")
sim1@params@gear_params$initial_effort <- 0
sim1@params@species_params$biomass_observed <- apply(rsts_total[,-10], 2, mean) * 1e3 # the survey values are in kg, converting to g

sim1 <- project(sim1@params)
plotBiomass(sim1)
plotBiomassObservedVsModelCustom(sim1)

params2 <- calibrateBiomassCustom(sim1@params)
plotBiomassObservedVsModelCustom(params2)


```

```{r, eval = F}
saveRDS(params2, file = "params/params11.rds")
```

Updating the observed biomass and scaling kappa up get us close to a good calibration already ;D

Improving the plot



```{r}

tempDat <- as.matrix(rsts_total[,-10]) *1e3

fqn <- apply(tempDat, 2, quantile, 0.25)
tqn <- apply(tempDat, 2, quantile, 0.75)

plot_dat <- plotBiomassObservedVsModelCustom(sim_optim, return_data = T)

plot_dat$fqn <- fqn
plot_dat$tqn <- tqn

ggplot(plot_dat, aes(x = observed, y = model,color = species)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = fqn, ymax = tqn)) +
  geom_abline(aes(intercept = 0, slope = 1), colour = "purple", 
              linetype = "dashed", size = 1.3) +
  coord_cartesian(ylim = range(plot_dat$model, plot_dat$observed)) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  scale_colour_manual(values = getColours(params2)[plot_dat$species]) + 
  ggrepel::geom_label_repel(aes(label = species), box.padding = 0.35, 
                            point.padding = 0.5, segment.color = "grey50", 
                            show.legend = FALSE, max.overlaps = Inf, seed = 42)






```


Need to rerun optim to get the biomass within the quantiles

## optimParallel

```{r}
params2 <- readRDS("params/params12.rds")
```


```{r}
# create set of params for the optimisation process
tic()
params_optim <- params2

vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@species_params$interaction_resource,
        params_optim@interaction)

params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim,  
                                             dat = params_optim@species_params$biomass_observed, method ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),rep(.01,9),rep(0.01,81)),
                                             upper= c(rep(15,9),rep(1,9),rep(.5,9),rep(0.8,81)),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_reboot4.RDS")
```


```{r}
optim_result <- readRDS("params/optim_result_reboot3.RDS")

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource<-optim_result$par[19:27]

interaction <- params_optim@interaction
interaction[] <- matrix(optim_result$par[28:108],nrow = 9) 
params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 1000)
```


```{r}
mizer::plotBiomass(sim_optim)
plotBiomassObservedVsModelCustom(sim_optim)
plotSpectra(sim_optim)
```

Looking good

```{r}

plotGrowthCurves(sim_optim, species_panel = T)
mizerHowTo::plotDiet2(sim_optim)
plotFeedingLevel(sim_optim)

sim_optim@params@species_params$interaction_resource
```

Feeding level is too high, linked to too fast growth. Need to reduce that. 

Let's tweak gamma with the shiny app

## tweaking gamma

```{r}
params <- sim_optim@params

newGamma <- mizerHowto::shiny_gamma(params)

sim_optim@params@species_params$gamma <- newGamma
params3 <- setParams(sim_optim@params)

sim3 <- project(params3)

plotBiomass(sim3)
plotBiomassObservedVsModelCustom(sim3)
plotGrowthCurves(sim3, species_panel = T)

params3@initial_n <- sim3@n[101,,]

params3 <- setParams(params3)
saveRDS(params3, file = "params/params12.rds")

```

Now need to optimise again with the new growth curves (doing optim on GEM48)


```{r}
params_optim <- readRDS("params/params12.rds")
optim_result <- readRDS("params/optim_result_reboot4.RDS")

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource<-optim_result$par[19:27]

interaction <- params_optim@interaction
interaction[] <- matrix(optim_result$par[28:108],nrow = 9) 
params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 1000)
```

```{r}
mizer::plotBiomass(sim_optim)
plotBiomassObservedVsModelCustom(sim_optim)
plotSpectra(sim_optim)

plotGrowthCurves(sim_optim, species_panel = T)
mizerHowTo::plotDiet2(sim_optim)
plotFeedingLevel(sim_optim)
```

```{r}

params4 <- sim_optim@params
params4@initial_n <- sim_optim@n[1001,,]
saveRDS(params4, file = "params/params13.rds")
```


# Fisheries calibration

The data is in ll_trends. Since I'm modelling the whole area, I can use the data straight like this. Effort is number of hooks and catch is in tonnes. I estimated the catchability already so should run straight the time series calibration.

```{r}

params4 <- readRDS("params/params13.rds")
```

Calibration setup. To match model to time series we are going to vary `R_max`, `erepro` and `catchability`.

```{r}

vary_df <- data.frame("name" = c("R_max","erepro","catchability"),
                      "length" = c(9,9,1),
                      "lower" = c(3,1e-7,1e-9), 
                      "upper" = c(15,.9,1e-2),
                      "slot" = c(rep("species_params",2),"gear_params"),
                      "unit" = c("log10",rep("linear",2)))

vary <- c(log10(params4@species_params$R_max),
          params4@species_params$erepro,
          params4@gear_params$catchability[1])

effort <-array(ll_trends$Effort, dim = c(length(ll_trends$Effort),1), 
               dimnames = list("time" = ll_trends$Year, "gear" = params4@gear_params$gear[1]))

time_series <- array(0, dim = c(dim(ll_trends)[1], dim(params4@species_params)[1]),
                     dimnames = list("time" = ll_trends$Year, "species" = params4@species_params$species))
time_series[,1] <- ll_trends$CatchT * 1e6 #conversion to g

```





```{r}
getErrorCustom(vary = vary,
               vary_df = vary_df, 
               params = params4, 
               data_type = "yield_observed", 
               effort = effort, 
               time_series = time_series)
```


it works, now need to adapt fastoptim and rebuild mizerhowto

```{r}

res <- mizerHowTo::fastOptim(params = params4,
                      vary = vary,
                      vary_df = vary_df,
                      errorFun = mizerHowTo::getErrorCustom, 
                      data_type = "yield_observed", 
                      effort = effort,
                      time_series = time_series) 




saveRDS(res, file = "params/optim_result_reboot5.RDS")

```




Template for fast optim
```{r}
library(mizerHowTo)

vary_df <- data.frame("name" = c("R_max","interaction_resource"),
                      "length" = c(12,12),
                      "lower" = c(3,.1), 
                      "upper" = c(15,.9),
                      "slot" = rep("species_params",2),
                      "unit" = c("log10","linear"))

vary <- c(log10(NS_params@species_params$R_max), rep(1,12))

sim <- project(NS_params)
NS_params@species_params$yield_observed <- getYield(sim)[101,]



getErrorCustom(vary = vary, vary_df = vary_df, params = NS_params, data_type = "yield_observed")

res <- fastOptim(params = NS_params, 
                 vary = vary,
                 vary_df = vary_df,
                 errorFun = getErrorCustom,
                 data_type = "yield_observed")
```

