---
title: doing the same calibration but better
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(terra)
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(therMizer)
library(parallel)
library(tictoc)
source("functions.R")

load("inputs/size_spec_inputs.RData")
```


# The data

I have two datasets. The biomass survey is RSTS data which is kg of species per area surveyed (effort column in km2). The longlines concerns the toothfish only and is in number of hooks (effort) and catches in tonnes. The data for the longline is for the whole area whereas the survey is per km2. The first step is to get everything to the same scale, in our case, the whole studied area.

# The studied area

Nicole sent an HIMI mask that I can use to get the total surface area

```{r}
HIMI_shape<-sf::st_read("HIMI_Mask/HIMI_EEZ_polygon/HIMI_EEZ_polygon.shp") # the shape file of the area
plot(HIMI_shape)
HIMI_area <- sf::st_area(HIMI_shape)
```

The surface area in km2 is

```{r}

total_area <- as.numeric(HIMI_area)/1e6
total_area

```

Now I need the fraction of surveyed area compared to the total area. This will allow me to model the shallows area with the juveniles within the bigger area.

The details of the area surveyed is in Area_surveyed.Rdata

```{r}

load("inputs/Area_surveyed.RData")

```


Now I can estimate the biomass for each species each year in the whole shallows. Hopefully, the sizes not caught by the survey but modelled by Mizer will match well what happens outside of the shallows


```{r}

rsts_total <- rsts_trends[,-(1:2)] * year_tot[3:16,]$tot_area / rsts_trends$Effort

rsts_total$year <- rsts_trends$Year

saveRDS(rsts_total, file = "inputs/rsts_total.rds")
```


Need to re-calibrate the model with these values. Are they more or less constant?

```{r}

plot_dat <- reshape2::melt(rsts_total)
plot_dat$year <- as.numeric(plot_dat$year)

ggplot(plot_dat) +
  geom_line(aes(x = year, y = value, color = variable)) +
  facet_wrap(~variable, scales = "free") +
  scale_y_continuous(trans = "log10")

```

Big no. Let's average the values for the calibration and plot standard deviation on the final plot

# Biomass calibration

The latest saved param object is params10


```{r}

sim1 <- readRDS("params/params10.RDS")
sim1@params@gear_params$initial_effort <- 0
sim1@params@species_params$biomass_observed <- apply(rsts_total[,-10], 2, mean) * 1e3 # the survey values are in kg, converting to g

sim1 <- project(sim1@params)
plotBiomass(sim1)
plotBiomassObservedVsModelCustom(sim1)

params2 <- calibrateBiomassCustom(sim1@params)
plotBiomassObservedVsModelCustom(params2)


```

```{r, eval = F}
saveRDS(params2, file = "params/params11.rds")
```

Updating the observed biomass and scaling kappa up get us close to a good calibration already ;D

Improving the plot



```{r}

tempDat <- as.matrix(rsts_total[,-10]) *1e3

fqn <- apply(tempDat, 2, quantile, 0.25)
tqn <- apply(tempDat, 2, quantile, 0.75)

plot_dat <- plotBiomassObservedVsModelCustom(sim_optim, return_data = T)

plot_dat$fqn <- fqn
plot_dat$tqn <- tqn

ggplot(plot_dat, aes(x = observed, y = model,color = species)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = fqn, ymax = tqn)) +
  geom_abline(aes(intercept = 0, slope = 1), colour = "purple", 
              linetype = "dashed", size = 1.3) +
  coord_cartesian(ylim = range(plot_dat$model, plot_dat$observed)) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  scale_colour_manual(values = getColours(params2)[plot_dat$species]) + 
  ggrepel::geom_label_repel(aes(label = species), box.padding = 0.35, 
                            point.padding = 0.5, segment.color = "grey50", 
                            show.legend = FALSE, max.overlaps = Inf, seed = 42)






```


Need to rerun optim to get the biomass within the quantiles

## optimParallel

```{r}
params2 <- readRDS("params/params12.rds")
```


```{r}
# create set of params for the optimisation process
tic()
params_optim <- params2

vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@species_params$interaction_resource,
        params_optim@interaction)

params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim,  
                                             dat = params_optim@species_params$biomass_observed, method ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),rep(.01,9),rep(0.01,81)),
                                             upper= c(rep(15,9),rep(1,9),rep(.5,9),rep(0.8,81)),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_reboot4.RDS")
```


```{r}
optim_result <- readRDS("params/optim_result_reboot3.RDS")

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource<-optim_result$par[19:27]

interaction <- params_optim@interaction
interaction[] <- matrix(optim_result$par[28:108],nrow = 9) 
params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 1000)
```


```{r}
mizer::plotBiomass(sim_optim)
plotBiomassObservedVsModelCustom(sim_optim)
plotSpectra(sim_optim)
```

Looking good

```{r}

plotGrowthCurves(sim_optim, species_panel = T)
mizerHowTo::plotDiet2(sim_optim)
plotFeedingLevel(sim_optim)

sim_optim@params@species_params$interaction_resource
```

Feeding level is too high, linked to too fast growth. Need to reduce that. 

Let's tweak gamma with the shiny app

## tweaking gamma

```{r}
params <- sim_optim@params

newGamma <- mizerHowto::shiny_gamma(params)

sim_optim@params@species_params$gamma <- newGamma
params3 <- setParams(sim_optim@params)

sim3 <- project(params3)

plotBiomass(sim3)
plotBiomassObservedVsModelCustom(sim3)
plotGrowthCurves(sim3, species_panel = T)

params3@initial_n <- sim3@n[101,,]

params3 <- setParams(params3)
saveRDS(params3, file = "params/params12.rds")

```

Now need to optimise again with the new growth curves (doing optim on GEM48)


```{r}
params_optim <- readRDS("params/params12.rds")
optim_result <- readRDS("params/optim_result_reboot4.RDS")

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource<-optim_result$par[19:27]

interaction <- params_optim@interaction
interaction[] <- matrix(optim_result$par[28:108],nrow = 9) 
params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 1000)
```

```{r}
mizer::plotBiomass(sim_optim)
plotBiomassObservedVsModelCustom(sim_optim)
plotSpectra(sim_optim)

plotGrowthCurves(sim_optim, species_panel = T)
mizerHowTo::plotDiet2(sim_optim)
plotFeedingLevel(sim_optim)
```

```{r}

params4 <- sim_optim@params
params4@initial_n <- sim_optim@n[1001,,]
saveRDS(params4, file = "params/params13.rds")
```


# Fisheries calibration

The data is in ll_trends. Since I'm modelling the whole area, I can use the data straight like this. Effort is number of hooks and catch is in tonnes. I estimated the catchability already so should run straight the time series calibration.



Template for fast optim
```{r}
library(mizerHowTo)

vary_df <- data.frame("name" = c("R_max","interaction_resource"),
                      "length" = c(12,12),
                      "lower" = c(3,.1), 
                      "upper" = c(15,.9),
                      "slot" = rep("species_params",2),
                      "unit" = c("log10","linear"))

vary <- c(log10(NS_params@species_params$R_max), rep(1,12))

sim <- project(NS_params)
NS_params@species_params$yield_observed <- getYield(sim)[101,]



getErrorCustom(vary = vary, vary_df = vary_df, params = NS_params, data_type = "yield_observed")

res <- fastOptim(params = NS_params, 
                 vary = vary,
                 vary_df = vary_df,
                 errorFun = getErrorCustom,
                 data_type = "yield_observed")
```



```{r}

params4 <- readRDS("params/params13.rds")
```

Calibration setup. To match model to time series we are going to vary `R_max`, `erepro` and `catchability`.

```{r}

vary_df <- data.frame("name" = c("R_max","erepro","catchability"),
                      "length" = c(9,9,1),
                      "lower" = c(3,1e-10,1e-15), 
                      "upper" = c(15,.9,1e-2),
                      "slot" = c(rep("species_params",2),"gear_params"),
                      "unit" = c("log10",rep("linear",2)))

vary <- c(log10(params4@species_params$R_max),
          params4@species_params$erepro,
          1e-9)

# effort <-array(ll_trends$Effort, dim = c(length(ll_trends$Effort),1), 
#                dimnames = list("time" = ll_trends$Year, "gear" = params4@gear_params$gear[1]))
# 
# time_series <- array(0, dim = c(dim(ll_trends)[1], dim(params4@species_params)[1]),
#                      dimnames = list("time" = ll_trends$Year, "species" = params4@species_params$species))
# time_series[,1] <- ll_trends$CatchT * 1e6 #conversion to g
# saveRDS(list(effort,time_series), file = "fishingInput.RDS")

myData <- readRDS("fishingInput.RDS")

effort <- myData[[1]]
time_series <- myData[[2]]

```


it works, now need to adapt fastoptim and rebuild mizerhowto

```{r}
tic()
res <- mizerHowTo::fastOptim(params = params4,
                             vary = vary,
                             vary_df = vary_df,
                             errorFun = mizerHowTo::getErrorCustom, 
                             data_type = "yield_observed", 
                             effort = effort,
                             time_series = time_series) 

toc()


saveRDS(res, file = "params/optim_result_reboot5.RDS")

```

Let's put things back in the param object and take a look!

```{r}
params_optim <- readRDS("params/params13.rds")
optim_result <- readRDS("params/optim_result_reboot5.RDS")

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1]<- optim_result$par[19]

saveRDS(params_optim, file = "params/params14.rds")

sim_optim <- project(params_optim, effort = effort)
```

Looping for more accuracy so looping optimparallel a few times


```{r}

vary_df <- data.frame("name" = c("R_max","erepro","catchability"),
                      "length" = c(9,9,1),
                      "lower" = c(3,1e-10,1e-15), 
                      "upper" = c(15,.9,1e-2),
                      "slot" = c(rep("species_params",2),"gear_params"),
                      "unit" = c("log10",rep("linear",2)))

vary <- c(log10(params_optim@species_params$R_max),
          params_optim@species_params$erepro,
          params_optim@gear_params$catchability[1])

myData <- readRDS("fishingInput.RDS")

effort <- myData[[1]]
time_series <- myData[[2]]

tic()
for(i in 1:10)
{
  
  res <- mizerHowTo::fastOptim(params = params_optim,
                               vary = vary,
                               vary_df = vary_df,
                               errorFun = mizerHowTo::getErrorCustom, 
                               data_type = "yield_observed", 
                               effort = effort,
                               time_series = time_series) 
  
  species_params(params_optim)$R_max<-10^optim_result$par[1:9]
  species_params(params_optim)$erepro<-optim_result$par[10:18]
  gear_params(params_optim)$catchability[1]<- optim_result$par[19]
  
  vary <- c(log10(params_optim@species_params$R_max),
            params_optim@species_params$erepro,
            params_optim@gear_params$catchability[1])
}

toc()

saveRDS(res, "params/optim_result_reboot6.RDS")


```

and let's see the result

```{r}
params_optim <- readRDS("params/params14.rds")
optim_result <- readRDS("params/optim_result_reboot6.RDS")
myData <- readRDS("fishingInput.RDS")
effort <- myData[[1]]
time_series <- myData[[2]]

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1]<- optim_result$par[19]

saveRDS(params_optim, file = "params/params15.rds")

sim_optim <- project(params_optim, effort = effort)
```



```{r}
plotSpectra(sim_optim)
plotBiomass(sim_optim)
plotYield(sim_optim)
plotGrowthCurves(sim_optim, species_panel = T)
mizerHowTo::plotDiet2(sim_optim)

```


meh

Comparing the yields

```{r}

plot_dat <- plotYield(sim_optim, return_data = T)
plot_dat$Species <- time_series[,1]
colnames(plot_dat)[2:3] <- c("modelled","observed")
plot_dat <- reshape2::melt(plot_dat, "Year")

ggplot(plot_dat) +
  geom_line(aes(x = Year, y = value, color = variable ))+
  scale_y_continuous(name = "Yield in g")


```


Fishing is obviously too high in the model, causing toothfish to collapse. Why was it selected by optimparallel?

Quick check of getError values

```{r}

vary_df <- data.frame("name" = c("catchability"),
                      "length" = 1,
                      "lower" = 1e-15, 
                      "upper" = 1e-5,
                      "slot" = "gear_params",
                      "unit" = "linear")

getErrorCustom(vary = 1e-9, 
               vary_df = vary_df, 
               params = sim_optim@params,
               data_type = "yield_observed",
               effort = effort,
               time_series = time_series
)

```

Simulations are not converging why?

```{r}

sim <- project(sim_optim@params, effort = 0)
plotBiomass(sim)

```

I think optimparallel changed rmax and erepro too much
Redoing calibration just with erepro and catchability


```{r}
params4 <- readRDS("params/params13.rds")

vary_df <- data.frame("name" = c("erepro","catchability"),
                      "length" = c(9,1),
                      "lower" = c(1e-10,1e-15), 
                      "upper" = c(.9,1e-2),
                      "slot" = c("species_params","gear_params"),
                      "unit" = c(rep("linear",2)))

vary <- c(
  params4@species_params$erepro,
  1e-9)

myData <- readRDS("fishingInput.RDS")

effort <- myData[[1]]
time_series <- myData[[2]]


tic()
res <- mizerHowTo::fastOptim(params = params4,
                             vary = vary,
                             vary_df = vary_df,
                             errorFun = mizerHowTo::getErrorCustom, 
                             data_type = "yield_observed", 
                             effort = effort,
                             time_series = time_series) 

toc()


saveRDS(res, file = "params/optim_result_reboot7.RDS")
res <- readRDS("params/optim_result_reboot7.RDS")

species_params(params4)$erepro<-res$par[1:9]
gear_params(params4)$catchability[1]<- res$par[10]

# saveRDS(params_optim, file = "params/params15.rds")
```


```{r}
sim_optim <- project(params4, effort = effort)
plotBiomass(sim_optim)

plot_dat <- plotYield(sim_optim, return_data = T)
plot_dat$Species <- time_series[,1]
colnames(plot_dat)[2:3] <- c("modelled","observed")
plot_dat <- reshape2::melt(plot_dat, "Year")

ggplot(plot_dat) +
  geom_line(aes(x = Year, y = value, color = variable))+
  scale_y_continuous(name = "Yield in g")

mizerHowTo::plotDiet2(sim_optim)

saveRDS(sim_optim, file = "params/params16.rds")

```


Yield looks way better and do not capture the sharp drop in 2015 so look promising :)
One of the prey is dying though

Now to fix M.cam, not really sure why it's going extinct... growth, feeding and mortality look okay

```{r}
sim_optim <- readRDS("params/params16.rds")

params5 <- calibrateBiomassCustom(sim_optim@params)
sim5 <- project(params5)
plotBiomass(sim5)
plotEnergyBudget(sim5)

```


Can also try to check the rdd vs rdi

```{r}

mizerHowTo::plotRdiVsRdd(sim5)

```

This shows that M Cam is not affected by rmax and just doesn't reproduce enough!

Let's play with wmat_25 and psi to tweak a better growth for M.cam, maybe it will fix the extinction

Current psi profile

```{r}

psi <- sim5@params@psi

plot_dat <- reshape2::melt(psi)

ggplot(plot_dat) +
  geom_line(aes(x = w, y = value, color = sp)) +
  scale_x_continuous(trans = "log10")


```

What is the current wmat_25 for the species? I'm going to assume that reducing it will make species grow and mature faster

```{r}

plot_dat <- plotGrowthCurves(sim5, species = "M.cam", return_data = T)
plot_dat$type <- "original"

sim5@params@species_params$w_mat25

sim6 <- sim5

sim6@params@species_params$w_mat25[5] <- sim5@params@species_params$w_mat25[5]*.85

params6 <- setParams(sim6@params)

sim6 <- project(params6)
plotBiomass(sim6)
res <- plotGrowthCurves(sim6, species = "M.cam", return_data = T)
res$type <- "updated"

plot_dat2 <- rbind(plot_dat, res)


ggplot(plot_dat2) +
  geom_line(aes(x = Age, y = value, color = type, linetype = Legend)) +
  geom_hline(yintercept = sim5@params@species_params$w_mat25[5], color = "red") +
  geom_hline(yintercept = sim6@params@species_params$w_mat25[5], color = "blue")


```

Can't manage to change the profile before ~age 5
Let's look at erepro instead

```{r}

mizerHowTo::shiny_erepro(sim5@params)

```


Bumping C.cam erepro to 0.3 should work


```{r}

params6 <- sim5@params
params6@species_params$erepro[5] <- .3
params6 <- setParams(params6)
params6 <- projectToSteady(params6)


sim6 <- project(params6)
plotBiomass(sim6)
plotBiomassObservedVsModelCustom(sim6)

sim6@params@initial_n <- sim6@n[dim(sim6@n)[1],,]


```

Checking if fisheries is still ok


```{r}
myData <- readRDS("fishingInput.RDS")
effort <- myData[[1]]
time_series <- myData[[2]]
sim6 <- project(sim6@params, effort = effort)
plot_dat <- plotYield(sim6, return_data = T)
plot_dat$Species <- time_series[,1]
colnames(plot_dat)[2:3] <- c("modelled","observed")
plot_dat <- reshape2::melt(plot_dat, "Year")

ggplot(plot_dat) +
  geom_line(aes(x = Year, y = value, color = variable))+
  scale_y_continuous(name = "Yield in g")

saveRDS(sim6, file = "params/params17.rds")
```


# Thermizer 

```{r}
sim6 <- readRDS("params/params17.rds")
params <- sim6@params
```

Setting up temperature range

```{r}

temp_min <- c(1.8,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.8,1.6) # fishbase
# species_params(params)$temp_min <- c(1.4,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.4,1.4) # adjusted so species survive historical temperatures

# temp_max <- c(8.8,2.9,2.5,2.8,1.4,2,2.3,3.1,3.9) #fishbase
temp_max <- c(8.8,2.9,2.5,2.8,2.4,2.4,2.3,3.1,3.9) # adjusted

```

Loading fismip data

```{r}
# temperature
ocean_temp_picontrol <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_PIcontrol.RDS")
ocean_temp_ssp1rcp26 <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_126.RDS")
ocean_temp_ssp5rcp85 <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_585.RDS")

# plankton
n_pp_array_picontrol <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_PIcontrol.RDS")
n_pp_array_ssp1rcp26 <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_126.RDS")
n_pp_array_ssp5rcp85 <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_585.RDS")
```

Set the scenarios

```{r}
params_IPSL_picontrol <- upgradeTherParams(params = params,
                                           temp_min = temp_min, temp_max = temp_max, 
                                           ocean_temp_array = ocean_temp_picontrol, 
                                           n_pp_array = n_pp_array_picontrol) 

params_IPSL_ssp1rcp26 <- upgradeTherParams(params = params,
                                           temp_min = temp_min, temp_max = temp_max, 
                                           ocean_temp_array = ocean_temp_ssp1rcp26, 
                                           n_pp_array = n_pp_array_ssp1rcp26)

params_IPSL_ssp5rcp85 <- upgradeTherParams(params = params,
                                           temp_min = temp_min, temp_max = temp_max, 
                                           ocean_temp_array = ocean_temp_ssp5rcp85, 
                                           n_pp_array = n_pp_array_ssp5rcp85)


```


Plot showing the temperature performance of the species vs what temperature there is and what to expect

```{r}
temperature <- params_IPSL_ssp1rcp26@other_params$other$ocean_temp[,1]

# encounter scalar value today and in the future
historical <- mean(temperature[40:60])
hist_scalar <- NULL
for(iTime in 1990:2009){
  temp <- scaled_temp_effect(params_IPSL_ssp1rcp26, t = iTime)[,1]
  hist_scalar <- cbind(hist_scalar,temp)
}
hist_scalar <- apply(hist_scalar,1,mean)

projection <- mean(temperature[124:144])
proj_scalar <- NULL
for(iTime in 2074:2093){
  temp <- scaled_temp_effect(params_IPSL_ssp1rcp26, t = iTime)[,1]
  proj_scalar <- cbind(proj_scalar,temp)
}
proj_scalar <- apply(proj_scalar,1,mean)

# metabolism scalar value today and in the future
scalarM <- NULL
for(iTemp in c(historical,projection)){
  temp_effect_metab_realms <- array(NA, dim = c(dim(other_params(params_IPSL_ssp1rcp26)$vertical_migration)), dimnames = c(dimnames(other_params(params_IPSL_ssp1rcp26)$vertical_migration)))

  # Using t+1 to avoid calling ocean_temp[0,] at the first time step
  # Looping through each realm
  nb_realms <- dim(other_params(params_IPSL_ssp1rcp26)$exposure)[1]
  for (r in seq(1, nb_realms, 1)) {
    temp_at_t <- iTemp

    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))

    # Arrhenius equation scaled to a value between 0 and 1
    temp_effect_metabolism_r <- (unscaled_temp_effect - species_params(params_IPSL_ssp1rcp26)$metab_min) / species_params(params_IPSL_ssp1rcp26)$metab_range

    # Set temperature effect to 0 if temperatures are outside thermal tolerance limits
    above_max <- temp_at_t > species_params(params_IPSL_ssp1rcp26)$temp_max
    below_min <- temp_at_t < species_params(params_IPSL_ssp1rcp26)$temp_min

    temp_effect_metabolism_r[above_max | below_min] = 0

    temp_effect_metab_realms[r,,] <- temp_effect_metabolism_r*other_params(params_IPSL_ssp1rcp26)$exposure[r,]*other_params(params_IPSL_ssp1rcp26)$vertical_migration[r,,]
  }

  temp_effect_metabolism <- colSums(temp_effect_metab_realms)
  scalarM <- cbind(scalarM, temp_effect_metabolism[,1])
}
  

plot_dat <- plotThermPerformance(params_IPSL_ssp1rcp26,resolution = .1,return_data = T)

plot_point <- data.frame("Species" = rep(params@species_params$species,2),
                        "temperature" = rep(c(rep(historical,length(params@species_params$species)),
                                          rep(projection,length(params@species_params$species))),2),
                        "scalar" = c(hist_scalar,proj_scalar,scalarM[,1],scalarM[,2]),
                        "Scenario" = rep(c(rep("historical",length(params@species_params$species)),
                                       rep("projection",length(params@species_params$species))),2),
                        "Type" = c(rep("Encounter",2*length(params@species_params$species)),
                                   rep("Metabolism",2*length(params@species_params$species))))


ggplot(plot_dat)+
  geom_line(aes(x = temperature, y = scalar, color = Type)) +
  geom_point(data = plot_point, aes(x = temperature, y = scalar, shape = Scenario, color = Scenario), size = 2) +
  scale_x_continuous("Temperature in C")+
  scale_y_continuous("Scalar value") +
  facet_wrap(~Species, scales = "free") +
  guides(color=guide_legend(nrow=2,byrow=TRUE), shape = guide_legend(nrow=2,byrow=TRUE)) +
  scale_color_manual(values = c("dodgerblue","darkgreen","green3","darkorange2"))

```

Version from 03thermizer.rmd looks better but know I jsut want to see which species temperature range I need to extend for them to survive at current temperature. Species changed: M.Cam and B.eat


Projections

```{r}

sim26 <- therProject(params_IPSL_ssp1rcp26)
plotBiomass(sim26)

```

For now everything dies ^^, slowly though. It would be good to have the species stable until 2009 at least.





