---
title: toothfish project
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(mizerHowTo)
library(parallel)
library(tictoc)
load("size_spec_inputs.RData")
```

### Helpful functions


```{r}
# function for calibration of Rmax

fastOptim <- function(params)
{

# create set of params for the optimisation process
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) # variable to explore
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError,params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

params_optim@species_params$R_max <- 10^optim_result$par 

sim_optim <- project(params_optim, t_max = 2000)

return(sim_optim)
}
```


```{r}
# function running tuneParams function in a row for a quick start to a calibration
fastCalib <- function(params, match = F)
{
params <- calibrateBiomass(params) # changes kappa and rmax
if(match) params <- matchBiomasses(params) # set rmax to inf and adjust erepro
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 1000)
return(sim)
}

```


```{r}
# removing effort = 1 so always using intial effort and removing /1e6 so everything is in grams

getError2 <- function (vary, params, dat, data_type = "catch", tol = 0.1, 
    timetorun = 10) 
{
    params@species_params$R_max[] <- 10^vary[1:length(params@species_params$R_max)]
    params <- setParams(params)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    sim <- project(params, t_max = timetorun, progress_bar = F)
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}


```


```{r}

getErrorTime <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]

  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred) - log(obs))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}


```


### Set of functions using size selectivity of the data

```{r}

# this function adds a lower boundary to selected size

plotBiomassObservedVsModelCustom <- function (object, species = NULL, ratio = FALSE, log_scale = TRUE, 
    return_data = FALSE, labels = TRUE, show_unobserved = FALSE) 
{
    if (is(object, "MizerSim")) {
        params = object@params
        n <- finalN(object)
    }
    else if (is(object, "MizerParams")) {
        params = object
        n <- initialN(params)
    }
    else {
        stop("You have not provided a valid mizerSim or mizerParams object.")
    }
    sp_params <- params@species_params
    species = valid_species_arg(object, species)
    if (length(species) == 0) 
        stop("No species selected, please fix.")
    row_select = match(species, sp_params$species)
    if (!"biomass_observed" %in% names(sp_params)) {
        stop("You have not provided values for the column 'biomass_observed' ", 
            "in the mizerParams/mizerSim object.")
    }
    else if (!is.numeric(sp_params$biomass_observed)) {
        stop("The column 'biomass_observed' in the mizerParams/mizerSim object", 
            " is not numeric, please fix.")
    }
    else {
        biomass_observed = sp_params$biomass_observed
    }
    
    cutoffLow <- sp_params$biomass_cutoffLow[row_select]
    if (is.null(cutoffLow)) {
        cutoffLow = rep(0, length(species))
    }
    else if (!is.numeric(cutoffLow)) {
        stop("params@species_params$biomass_cutoffLow is not numeric, \",\n                 \"please fix.")
    }
    cutoffLow[is.na(cutoffLow)] <- 0
    
    cutoffHigh <- sp_params$biomass_cutoffHigh[row_select]
    if (is.null(cutoffHigh)) {
        cutoffHigh = rep(0, length(species))
    }
    else if (!is.numeric(cutoffHigh)) {
        stop("params@species_params$biomass_cutoffHigh is not numeric, \",\n                 \"please fix.")
    }
    cutoffHigh[is.na(cutoffHigh)] <- 0
    
    sim_biomass = rep(0, length(species))
    for (j in 1:length(species)) {
        sim_biomass[j] = sum((n[row_select[j], ] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    dummy = data.frame(species = species, model = sim_biomass, 
        observed = biomass_observed[row_select]) %>% mutate(species = factor(species, 
        levels = species), is_observed = !is.na(observed) & observed > 
        0, observed = case_when(is_observed ~ observed, !is_observed ~ 
        model), ratio = model/observed)
    if (sum(dummy$is_observed) == 0) {
        cat(paste("There are no observed biomasses to compare to model,", 
            "only plotting model biomasses.", sep = "\n"))
    }
    if (!show_unobserved) {
        dummy <- filter(dummy, is_observed)
    }
    if (return_data == TRUE) 
        return(dummy)
    tre <- round(sum(abs(1 - dummy$ratio)), digits = 3)
    caption <- paste0("Total relative error = ", tre)
    if (any(!dummy$is_observed)) {
        caption <- paste(caption, "\n Open circles represent species without biomass observation.")
    }
    if (ratio == FALSE) {
        gg <- ggplot(data = dummy, aes(x = observed, y = model, 
            colour = species, shape = is_observed)) + geom_abline(aes(intercept = 0, 
            slope = 1), colour = "purple", linetype = "dashed", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass [g]") + 
            coord_cartesian(ylim = range(dummy$model, dummy$observed))
    }
    else {
        gg <- ggplot(data = dummy, aes(x = observed, y = ratio, 
            colour = species, shape = is_observed)) + geom_hline(aes(yintercept = 1), 
            linetype = "dashed", colour = "purple", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass / observed biomass") + 
            coord_cartesian(ylim = range(dummy$ratio))
    }
    gg <- gg + labs(x = "observed biomass [g]", caption = caption) + 
        scale_colour_manual(values = getColours(params)[dummy$species]) + 
        scale_shape_manual(values = c(`TRUE` = 19, `FALSE` = 1)) + 
        guides(shape = "none")
    if (log_scale == TRUE & ratio == FALSE) {
        gg = gg + scale_x_log10() + scale_y_log10()
    }
    if (log_scale == TRUE & ratio == TRUE) {
        gg = gg + scale_x_log10()
    }
    if (labels == TRUE) {
        gg = gg + ggrepel::geom_label_repel(aes(label = species), 
            box.padding = 0.35, point.padding = 0.5, segment.color = "grey50", 
            show.legend = FALSE, max.overlaps = Inf, seed = 42)
    }
    gg
}
```


```{r}

# adapting cutoff here too

calibrateBiomassCustom <- function (params) 
{
  if ((!("biomass_observed" %in% names(params@species_params))) || 
      all(is.na(params@species_params$biomass_observed))) {
    return(params)
  }
  no_sp <- nrow(params@species_params)
  
  cutoffLow <- params@species_params$biomass_cutoffLow
  if (is.null(cutoffLow)) 
    cutoffLow <- rep(0, no_sp)
  cutoffLow[is.na(cutoffLow)] <- 0
  
  cutoffHigh <- params@species_params$biomass_cutoffHigh
  if (is.null(cutoffHigh)) 
    cutoffHigh <- rep(0, no_sp)
  cutoffHigh[is.na(cutoffHigh)] <- 0
  
  observed <- params@species_params$biomass_observed
  observed_total <- sum(observed, na.rm = TRUE)
  sp_observed <- which(!is.na(observed))
  model_total <- 0
  for (sp_idx in sp_observed) {
    model_total <- model_total + sum((params@initial_n[sp_idx, 
    ] * params@w * params@dw)[params@w >= cutoffLow[sp_idx] & cutoffHigh[sp_idx] >= params@w])
  }
  scaleModel(params, factor = observed_total/model_total)
}


```

```{r}

# same as above

matchBiomassCustom <- function (params, species = NULL) 
{
    if (!("biomass_observed" %in% names(params@species_params))) {
        return(params)
    }
    species <- valid_species_arg(params, species = species, return.logical = TRUE) & 
        !is.na(params@species_params$biomass_observed) & params@species_params$biomass_observed > 
        0
    for (sp in (1:nrow(params@species_params))[species]) {
        cutoffLow <- params@species_params$biomass_cutoffLow[[sp]]
        if (is.null(cutoffLow) || is.na(cutoffLow)) {
            cutoffLow <- 0
        }
                    cutoffHigh <- params@species_params$biomass_cutoffHigh[[sp]]
        if (is.null(cutoffHigh) || is.na(cutoffHigh)) {
            cutoffHigh <- 0
        }
        
        total <- sum((params@initial_n[sp, ] * params@w * params@dw)[params@w >= cutoffLow & cutoffHigh >= params@w])
        factor <- params@species_params$biomass_observed[[sp]]/total
        params@initial_n[sp, ] <- params@initial_n[sp, ] * factor
    }
    setBevertonHolt(params)
}



```


```{r}
getErrorCustom <- function(vary, params, species, dat, tol = 0.001, 
    timetorun = 10)
{
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]

  params <- setParams(params)

  interaction <- params@interaction
  interaction[species,] <- vary[19:27]
  
  params <- setInteraction(params,interaction)

    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    sim_biomass = rep(0, length(params@species_params$species))
    
        cutoffLow <- params@species_params$biomass_cutoffLow
    if (is.null(cutoffLow)) 
        cutoffLow <- rep(0, no_sp)
    cutoffLow[is.na(cutoffLow)] <- 0
    
        cutoffHigh <- params@species_params$biomass_cutoffHigh
    if (is.null(cutoffHigh)) 
        cutoffHigh <- rep(0, no_sp)
    cutoffHigh[is.na(cutoffHigh)] <- 0
        
    for (j in 1:length(sim_biomass)) {
        sim_biomass[j] = sum((sim@n[dim(sim@n)[1],j,] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    

    pred <- log(sim_biomass)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```


For setup, run all the blocks above. Do not knit.
Latest simulation calibrated is sim17 (saved as params/params7.RDS)

### Calibration

The following param object is a Mizer params object assembled from different dataset from the toothfish team. So it's either empirical data or default mizer value. The interaction matrix is approximated from a cooccurence matrix of the survey, however the values are rather low so they've been multiplied by 2.

```{r}
params <- readRDS("params/params0.RDS")
```

What does it look like?

```{r}
sim <- project(params)

plotCalibration(sim)
```


Let's use the functions from tuneParams first!

```{r}
params_steady <- steady(params)
# running it once, did not converge and some erepro are over 1

params_steady <- calibrateBiomass(params_steady) 
plotBiomassObservedVsModel(params_steady)
# looks like chaos, but we just matched total biomass of species, not the spectrum themselves

params_steady <- matchBiomasses(params_steady)
plotSpectra(params_steady)
plotBiomassObservedVsModel(params_steady)
# matching biomasses but have to recalculate steady state

params_steady <- steady(params_steady)
# converged this time
plotSpectra(params_steady, power = 2)

sim <- project(params_steady, t_max = 1000)
plotBiomass(sim)
# still not steady for some reasons
```


trying to find steady state from end of previous sim


```{r}
initialN(params_steady) <- sim@n[dim(sim@n)[1],,]
params_steady2 <- steady(params_steady,t_max = 1000)
sim2 <- project(params_steady2, t_max = 1000)
plotBiomass(sim2)
```


This is not going to work as long as erepro > 1

Rematching biomass

```{r}

sim3 <- fastCalib(sim2@params)
plotBiomass(sim3)

```

erepro are < 1 now but C.rhi goes extinct from ~300 years 

```{r}
plotBiomassObservedVsModel(sim3)
```


going to setup Rmax to the intial abundance of sim4 where everything coexist

```{r}
sim4 <- sim3
sim4@params@species_params$R_max <- apply(sim4@params@initial_n,1,sum)
sim4 <- project(sim4@params, t_max = 1000)
plotBiomass(sim4)
```

Doesn't work at all
```{r}
params_steady4 <- steady(sim4@params, tol = 0.001)
sim5 <- project(params_steady4, t_max = 1000)
plotBiomass(sim5)
```


now we are back to a similar result of sim3, one species still go extinct, after 300 years of coexistence

```{r}
plotGrowthCurves(sim5, species_panel = T)
plotDiet2(sim5)
```

This looks pretty good though

C.rhi is eaten a lot, maybe that's why
```{r}

sim5@params@interaction
# going to reduce the exposure of C.rhi
sim5@params@interaction[,3] <- sim5@params@interaction[,3]*0.75

#lets start from the beginning

sim6 <- fastCalib(sim5@params)
plotBiomass(sim6)

```

 this is not working
 
 let's do without matching biomass first, any coexistence is welcome
 
```{r}
params_steady6 <- steady(sim5@params, tol = 0.001)
sim6 <- project(params_steady6, t_max = 500)
plotBiomass(sim6)
```

species are declining, no convergence is found

```{r}
plotDiet2(sim6)
plotGrowthCurves(sim6, species_panel = T)
```

somehow it looks like species feed more on C.rhi even after I scaled down the column in the interaction matrix to reduce predation pressure on them


```{r}
plotPredMort(sim6)

```
 
Reasonable?

Saving it as intermediate step

```{r, eval = F}

saveRDS(sim6, file = "params/params1.RDS")
sim6 <- readRDS("params/params1.RDS")

```



Let's try a wide range of rmax

```{r}

sim7 <- fastOptim(sim6@params)
plotBiomass(sim7)
```


back to the beginning, C rhi is still going extinct, but with another species now


```{r}
sim6@params@species_params$beta
```

Default mizer beta are too low probably

```{r}

sim7 <- sim6

sim7@params@species_params$beta[3:9] <- 100

sim7 <- fastCalib(sim7@params)

plotBiomass(sim7)
```

Says there is convergence but everything collapses.

```{r}
plotDiet2(sim7)

```

Diets are not too bad even with the increase in beta 

Rmax values have been reset by the matchBiomass function, let's recalculate them


```{r}

sim7@params@species_params$R_max <- sim6@params@species_params$R_max

sim8 <- fastOptim(sim7@params)
```


```{r}

plotBiomass(sim8)
```

Still not stable, let's try to calibrate again

```{r}

sim9 <- fastCalib(sim8@params)
plotBiomass(sim9)

```

Matching the biomass always put Rmax to inf, what does that mean? They don't feed enough?


```{r}

plotBiomassObservedVsModel(sim9@params)

```


What if I reset Rmax with the default behavior


```{r}

sim9@params@species_params$R_max <- resource_params(sim9@params)$kappa * species_params(sim9@params)$w_inf^-1

sim9 <- project(sim9@params, t_max = 3000)

plotBiomass(sim9)

```

It does seem stable now, what about the biomass?

Saving that one in any cases

```{r, eval = F}

saveRDS(sim9, file = "params/params2.RDS")
sim9 <- readRDS("params/params2.RDS")

```



```{r}
plotBiomassObservedVsModel(sim9)
```

meh


```{r}

sim10 <- fastCalib(sim9@params)
plotBiomass(sim10)

```

meh

Can't have coexistence and right biomass so what's wrong? parameters?


Let's keep sim9 for now

```{r}

plotCalibration(sim9)
plotDiet2(sim9)
plotGrowthCurves(sim9, species_panel = T)
plotPredMort(sim9)

# let's work on these params

params <- sim9@params


```

Now let's calibrate it to the fishing time series following mizerHowTo3

There is only one type of gear, trawls, which catches everything (but mostly toothfish)

```{r}
load("size_spec_inputs.RData")

plot_dat <- rsts_trends[,-2]
plot_dat <- reshape2::melt(plot_dat, "Year")
plot_dat$Year <- as.numeric(plot_dat$Year)

ggplot(plot_dat) +
  geom_line(aes(x = Year, y = value, color = variable)) +
  scale_y_continuous(trans = "log10")
```


How to estimate a catchability value per species from this?

If we assume that biomass caught is effort * catchability * pristine biomass and I know the biomass caught and the effort, I'll guess the catchability. It means I need to use the unfished biomass from sim9 though, which I know is not properly calibrated, but no choice for now.

I can do that per year and estimate and then get an averaged catchability

```{r}

# unfished biomass
biom0 <- getBiomass(sim9)[dim(sim9@n)[1],] # quite annoying that D.ele doesn't have the highest biomass ;(

biomF <- rsts_trends

catchSim <- data.frame("year" = rsts_trends$Year)
for(icol in 3:11){

catchSp <- biomF[,icol] / (biomF$Effort * biom0[c(icol-2)])
catchSim <- cbind(catchSim,catchSp)
}
```


Let's do an average catchability


```{r}

catchabilityAvg <- apply(catchSim[,-1],2,mean)

catchArray <- matrix(catchabilityAvg, nrow = 1, dimnames = list("gear" = "knife_edge_gear", "sp" = params@species_params$species))

mizer::catchability(params) <- catchArray
# line above updates catchability slot but not gear nor species params slot

params <- setParams(params) # that one doesn't either. why so complicated !!!

params@species_params$catchability <- catchabilityAvg

params <- setParams(params)

params <- setFishing(params, catchability = catchArray)

```

Can't set up the fucking catchability

And I'm looking at the survey anyway, which I just used to calibrate the biomass so not sure if useful or not...


Now using longlines fisheries

```{r}
plot_dat <- reshape2::melt(ll_trends, "Year")

ggplot(plot_dat, aes(x = Year, y = value)) +
  geom_point() +
  facet_wrap(~variable, nrow = 2, scales = "free")
```



Let's do the same as above but with toothfish only

Average catchability

```{r}

# unfished biomass
biom0 <- getBiomass(sim9)[dim(sim9@n)[1],] # quite annoying that D.ele doesn't have the highest biomass ;(

biomF <- ll_trends

biomF$CatchT <- biomF$CatchT * 1e6 # converting tonnes to g

biomF$biom0 <- biom0[1]

```

Not even gonna try, the catch is way higher than what there actually is...

Need a proper calibration first


```{r}

plotBiomassObservedVsModel(sim9)

```

I need to look at predation, I feel that's where the problem is. So interaction matrix, beta and kappa

```{r}

sim10 <- sim9 


params <- calibrateBiomass(sim10@params) 
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 500)
plotBiomass(sim)
plotDiet2(sim)

plotBiomassObservedVsModel(sim)

```

Doing this little test shows that the function matchbiomass is a pain as it removes rmax and break the ecosystem. Steady is also bullshiting as it says one species needs erepro >1 and doesn't get steady but it actually does. punk.
However, D.ele is still too low

Biomass for the survey is in fucking kilos and I have grams ahahahahahhahahahahahahahahahaha
Maybe it will change something


```{r}
sim9@params@species_params$biomass_observed <- sim9@params@species_params$biomass_observed *1000
```


```{r}

sim10 <- fastCalib(sim9@params)

plotBiomass(sim10)
plotBiomassObservedVsModel(sim10)
plotDiet2(sim10)
plotSpectra(sim10)
plotGrowthCurves(sim10,species_panel = T)
```

Not much change except that kappa is higher now and the diets are more centered on plankton
But know I should be able to catch with long lines


```{r}

# unfished biomass
biom0 <- getBiomass(sim10)[dim(sim10@n)[1],] # quite annoying that D.ele doesn't have the highest biomass ;(

biomF <- ll_trends

biomF$CatchT <- biomF$CatchT * 1e6 # converting tonnes to g

biomF$biom0 <- biom0[1]

```

Catches are still too high
How do I increase my biomass without collapsing the ecosystem?
Need a plot comparing reproduction with and without rmax to see which species are impacted the most

I need to get biomass high enough meaning better Rmax



```{r}


sim11 <- fastOptim(sim10@params)
plotBiomass(sim11)
plotBiomassObservedVsModel(sim11)
```

Looks good, could try to fix this skate dying

```{r}

saveRDS(sim11, file = "params/params3.RDS")
sim11 <- readRDS("params/params3.RDS")

```

So what's wrong with you Bmur?

```{r}
plotGrowthCurves(sim11, species_panel = T)
plotDiet2(sim11)
plotPredMort(sim11)

```

Do you get eaten to much by meanie Dele?

```{r}
sim11@params@interaction[1,9] <- .5
params11 <- setParams(sim11@params)
sim12 <- project(params11, t_max = 500)
plotBiomass(sim12)

```

Bam EZ

```{r}

plotBiomassObservedVsModel(sim12)

```

A bit too much maybe ^^'


```{r}
sim11@params@interaction[1,9] <- .85
params11 <- setParams(sim11@params)
sim12 <- project(params11, t_max = 500)
plotBiomass(sim12)
plotBiomassObservedVsModel(sim12)

```

This seem to work great! Maybe the best way would be to tweak the interaction matrix to match the biomass?
How do I do that?

Just noticed something, the getError function I'm using uses catch as default argument, and therefore calculate Rmax based on the yield but I don't have any fisheries enabled.......... Corrected now


Let's try that here. First need a geterror function using a column of the interaction matrix


```{r}

vary <- rep(.5,9)
species <- params@species_params$species[1]

getError2 <- function(vary, params, species, dat, data_type = "catch", tol = 0.001, 
    timetorun = 10)
{
  interaction <- params@interaction
  interaction[species,] <- vary
  
  params <- setInteraction(params,interaction)

    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]/1e+06
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}


fastOptim2 <- function(params)
{

# create set of params for the optimisation process
params_optim <- params
vary <-  rep(.5,9) # variable to explore
species = "D.ele"
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError2,params=params_optim, species = species, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method ="L-BFGS-B", lower=c(rep(0,dim(params_optim@species_params)[1])), upper= c(rep(1,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)


  interaction <- params_optim@interaction
  interaction[species,] <- optim_results$par
  params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 2000)

return(sim_optim)
}



```


YOLO


```{r}

sim13 <- fastOptim2(sim12@params)

```

weird that it settles somewhere where it doesn't coexist. but now at least D.ele has the highest biomass


```{r}

sim13@params@interaction
plotPredMort(sim13)
plotDiet2(sim13)
plotBiomassObservedVsModel(sim13)

```

```{r}

saveRDS(sim13, file = "params/params4.RDS")
sim13 <- readRDS("params/params4.RDS")

```


Let's tweak the interaction manually


```{r}

interaction <- sim13@params@interaction
interaction[1,8] <- .9 * interaction[1,8]
interaction[1,9] <- .8 * interaction[1,9]
params <- setInteraction(sim13@params,interaction)

sim14 <- project(params, t_max = 2000)

```


```{r}

plotBiomass(sim14)
plotBiomassObservedVsModel(sim14)

```

Not bad at all :D


```{r}
saveRDS(sim14, file = "params/params5.RDS")
sim14 <- readRDS("params/params5.RDS")
```



Looking at the trawl vs longlines

```{r}

trawl <- rsts_trends[,1:3]
lline <- ll_trends[1:14,]

plot_dat <- data.frame("Year" = trawl[,1], "Trawl" = trawl[,3]/1000, "Longline" = lline[,3])
colnames(plot_dat)[2:3] <- c("Trawl","Longline")

ggplot(plot_dat) +
  geom_point(aes(x = Year, y = Trawl), color = "black") +
  geom_point(aes(x = Year, y = Longline), color = "red")


```

The biomass caught by the longlines is way higher than what's reported by the survey. It's because they are not in the same area and there is way more stuff, or at least toothfish in the depth. So whatever calibration I'm doing with the survey is not really useful I guess...

Maybe I should try to calibrate with the longlines only and the rest of the ecosystem will follow? And somehow keep the same proportion of other species, but with the biomass of the longlines? :/

Restarting the calibration from scratch, but using only toothfish biomass. I need a catchability to work from fisheries right? So it's not gonna work.

Can I assume that the deep ecosystem is just a big version of the shallows? Guessing that's the easiest to do at the moment. So the survey gives me the proportions of biomasses and I can scale it up so that the long lines do not fish everything

At the moment, params5 has ~5e10 g of D.elem so 5e4 Tonnes. Longlines catch up to 4e3 Tonnes so maybe that's enough

But before that, I converted twice kilos to grams so params5 has 1000 times more biomass than what it's supposed to have ....


```{r}

sim14@params@species_params$biomass_observed <- sim14@params@species_params$biomass_observed / 1000

sim15 <- fastCalib(sim14@params)

plotBiomass(sim15)
plotBiomassObservedVsModel(sim15)
plotDiet2(sim15)
plotSpectra(sim15)
plotGrowthCurves(sim15,species_panel = T)

```

sim15 doesn't work, going to try to just scale down the model. What a mess



```{r}

params15 <- mizer::scaleModel(sim14@params,factor = 1/1000)

sim15 <- project(params15, t_max = 500)

plotBiomass(sim15)
plotBiomassObservedVsModel(sim15)

```

Good
Now the ecosystem is truly calibrated to the survey. Let's fish it.

First thing, there is not enough D.ele to fish

If I knew the catchability of the long lines, I could deduce how much D.ele there is in the depth, but I don't. I just know the quantity know. Don't even know if the stock is collapsing or not.

Ok, let's assume, since the survey is relatively stable from 2004 to 2016, that the toothfish stock is stable in the depth too. So the long lines should not affect to much the stock of D.ele until 2016. Could try to scale up the ecosystem until fisheries do not collapse everything. Then we can see the effect of increasing the effort so much post 2016.

Since I do not have catchability, I'll assume it's 1.

Let's set up longlines for toothfish only for now. Will add bycatch later.
The graphs sent by Nicole shows that the smallest toothfish caught is 500mm long. However I'm simulation knife edge so I'm going to take 600mm minimum as there are not many 500mm toothfish caught. Best would be to do a sigmoid but I do not have the graph's data.

For toothfish, 60cm is 0.00603 * 60 ^ 3.18 = 2721.688g. Going to set up the min at that. It is way lower than w_mat though which is 95cm. If this is true, the longlines fish toothfish before they are mature. What a bunch of idiots.


```{r}
params <- sim16@params

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(1,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll

```




```{r}

sim16 <- project(params, t_max = 500, effort = 0)
plotSpectra(sim16, power = 2)
plotBiomass(sim16)
plotlySpectra(sim16, power = 2)
plotGrowthCurves(sim16, species = "D.ele")
```

```{r}
saveRDS(sim16, file = "params/params6.RDS")
sim16 <- readRDS("params/params6.RDS")

```



```{r}

sim17 <- project(sim16@params, t_max = 500, effort = 1)
plotBiomass(sim17)

```

Toothfish doesn't like the fisheries


```{r}

plotFMort(sim17)
plotPredMort(sim17)
plotSpectra(sim17)
```

Toothfish grow big and has a slow growth and yet they fish just small young individuals? WTF

Issue: how can the slow growth counterbalance the fisheries mortality?

```{r}

plot_dat <- plotEnergyBudget(sim16, species = "D.ele", return_data = T)
params <- sim16@params
 params <- setColours(params, list(Growth = "#F8766D", 
        Income = "#7CAE00", `Metabolic loss` = "#00BFCFC4", 
        Reproduction = "#C77CFF"))
xtrans <- "log10"
 
pl <- plotDataFrame(plot_dat[[1]], params, style = "area", 
        xlab = "Size [g]", ylab = "Rate [g/year]", 
        xtrans = xtrans, wrap_var = "Species", wrap_scale = "free")

    pl <- pl + geom_vline(data = plot_dat[[2]], aes(xintercept = w_mat, 
        group = Species), linetype = "dotted") + 
      geom_vline(data = plot_dat[[2]], 
        aes(xintercept = w_inf, group = Species), linetype = "dotted") + 
        geom_text(data = plot_dat[[2]], aes(x = w_mat, y = y_coord * 
            0.2, label = "\nMaturity"), angle = 90) + 
      geom_text(data = plot_dat[[2]], 
        aes(x = w_inf, y = y_coord * 0.2, label = "\nMaximum"), 
        angle = 90)+
      geom_vline(xintercept = 2722, color = "red")


```


Above graph shows that in this fisheries conditions, toothfish cannot reproduce. But the modelled growth is already slightly faster than the vonB one...


In the meantime let's recreate the longlines at a km2 scale

```{r}

ll_area_yr <- read_csv("ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)

ll_avg <- mean(ll_km2$catchT) * 1e6 # in grams
```

what does it look like?


```{r}

ggplot(ll_km2) +
  geom_line(aes(x = year, y = catchT), color = "red") +
  geom_line(aes(x = year, y = effort/1000), color = "black")

```


Redoing the fisheries and matching to time series


```{r}

params <- sim16@params

params@species_params$yield_observed <- c(ll_avg,rep(0,8))

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(0.00001,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll
initial_effort(params) <- 1 # somehow doesn't work with gear_params

sim16 <- project(params, t_max = 500)
plotBiomass(sim16)
plotFMort(sim16)

plotYieldObservedVsModel(sim16, species = "D.ele")
```



```{r}

params17 <- sim16@params
params17@initial_n <- sim16@n[dim(sim16@n)[1],,]
params17@initial_n_pp <- sim16@n_pp[dim(sim16@n_pp)[1],]

params17 <- setParams(params17)

sim17 <- project(params17)

plotBiomass(sim17)

saveRDS(sim17, file = "params/params7.RDS")

params <- sim17@params

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(1e-5,rep(0,8)),
               initial_effort = mean(ll_km2$effort))

gear_params(params) <- gear_ll

sim18 <- project(params, effort = params@gear_params$initial_effort[1], t_max = 500)

plotBiomass(sim18)

plotFMort(sim18)





```

if does not work do the optim on time averaged first


At least it works, nevertheless the low catchability

Let's run it through optim varying erepro, rmax and catchability


```{r}
tic()
dat <- ll_km2

  effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

# create set of params for the optimisation process
params_optim <- sim17@params
vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@gear_params$catchability[1])
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorTime,params=params_optim, dat = ll_km2, effort = effort,
                                             method   ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),0.000001),upper= c(rep(15,9),rep(1,9),1),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_timeseries.RDS")
optim_result <- readRDS("params/optim_result_timeseries.RDS")
#put these new vals intospecies_params and go back to the top of this page to re-check the calibration 
species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1] <-optim_result$par[19]
#re-run time-varying effort model tthough time with new erepro
sim_optim <- project(params_optim, effort = effort)




```


```{r}
plotBiomass(sim_optim)
plotFMort(sim_optim)
plotYield(sim_optim)
```

This is not working



Starting back from sim 17 but with the size data of toothfish of the survey now. So I can get more biomass for toothfish as only the juveniles come from the survey.

Two ways > either change the plotobservedbiomass to take into account only size class below average max size for toothfish or extrapolate adult biomass using a size spectra I guess? Not sure about second one though


Getting a rough idea of the average maximum size per year for toothfish
```{r}
sim17 <- readRDS("params/params7.RDS")

sizeDat <- get(load("RSTS_length_SizeSpec.Rdata"))

# To get a snapshot of the data, need to use Number of fish column

sizeFreq <- NULL
for(iYear in unique(sizeDat$Year)){
  tempDat <- filter(sizeDat, Year == iYear)
  for(iRow in 1:dim(tempDat)[1]){
    TL <- rep(tempDat$TL[iRow],tempDat$NumFish[iRow])
    sizeFreq <- rbind(sizeFreq, data.frame("TL" = TL, Year = iYear))
  }
}

# Any outliers?
sizeFreq$Year <- as.factor(sizeFreq$Year)

ggplot(filter(sizeFreq, Year %in% c(1999, 2001, 2002))) +
  geom_density(aes(x=TL, color = Year, group = Year))
```

Need to decide a maximum size and minimum size from this mess. Best would be one value for all years

Let's calculate the z score across all years

```{r}

sizeFreq$zscore <- (sizeFreq$TL - mean(sizeFreq$TL)) / sd(sizeFreq$TL)

summary(sizeFreq$zscore)

# the kde function gives the kernel density estimate of the data which is the equivalent of what geom_density does but not smoothed
res <- ks::kde(sizeFreq$zscore)
kde <- data.frame("eval" = res$eval.points, "density" = res$estimate)

ggplot(filter(sizeFreq)) +
  geom_density(aes(x=zscore)) +
  geom_vline(xintercept = -0.6842, linetype = "dashed") +
  geom_vline(xintercept = 0.5740, linetype = "dashed") +
  geom_line(data = kde, aes(x = eval, y = density), color = "red")

# going to take the 95% highest values of the kde to remove the smallest and biggest individuals

orderedKde <- kde[order(kde$density, decreasing = T),]

threshold <- round(.95 * dim(orderedKde)[1])

truncatedKDE <- orderedKde[1:threshold,]

summary(truncatedKDE$eval)

# don't know where I am going since I cannot get back to the zscore anyway...

# going to take the 95% lowest zscore instead


freqOrder <- sizeFreq[order(sizeFreq$zscore, decreasing = F),]

freqTrunc <- freqOrder[1:round(.95 * dim(freqOrder)[1]),]

summary(freqTrunc$TL)
summary(sizeFreq$TL)
```

Apparently, most of the outliers were the biggest individuals so going to assume that the survey data caught all individuals between 264mm and 690.5mm and start the size spectrum from here. Assuming that for all species.


Converted to grams

```{r}

#species_df$w_inf <- species_df$a * species_df$l_inf^species_df$b

sim17@params@species_params[1,]$a * 26.4^sim17@params@species_params[1,]$b
sim17@params@species_params[1,]$a * 69.05^sim17@params@species_params[1,]$b

```

Between 200 g and 4254.6 g

The closest size bins to that are

```{r}

res <- sim17@params@w[which(sim17@params@w >= 100)]

which(sim17@params@w == res[4])
which(sim17@params@w == res[20])

```

Ok so there is a size bin at 199g (65th) and one at 4209g(81th) so I'm going to take these ones.
Taking the range 65 to 80 (cause 81 starts at 4209g)

Also we could use the cutoff argument in plotbiomassvsobserved ahaha. Just need to add a high cutoff (for now only lower sizes are cut off)


```{r}

sim17@params@species_params$biomass_cutoffLow <- sim17@params@w[65] 
sim17@params@species_params$biomass_cutoffHigh <- sim17@params@w[81]

```


Now let's check that the new truncated functions work correctly
At the moment sim17 is on an equilibrium state calibrated without size selection


```{r}

biom1 <- getBiomass(sim17)[100,]

params18 <- calibrateBiomassCustom(sim17@params)

getBiomass(params18)

plotSpectra(params18)
plotSpectra(sim17)

initBiom <- params18@initial_n * params18@w * params18@dw

initBiom <- sweep(params18@initial_n, 2, params18@w * params18@dw, "*")

apply(initBiom[,65:81],1,sum)

params18@species_params$biomass_observed

params18 <- matchBiomassCustom(params18)

initBiom <- params18@initial_n * params18@w * params18@dw

initBiom <- sweep(params18@initial_n, 2, params18@w * params18@dw, "*")

apply(initBiom[,65:81],1,sum)

plotBiomassObservedVsModelCustom(params18)

plotBiomassObservedVsModel(params18) # this shows how much biomass is added thanks to size selectivity of the raw data

```


Ok so the biomass matches. Let's start again calibrating from params 18
L.squ has a erepro slighly higher than 1, need to remove that


```{r}

params18@species_params

params18 <- calibrateBiomassCustom(params18)
plotBiomassObservedVsModelCustom(params18)

params <- steady(params18)

# lots of erepro need to be higher than one here, what if I just scale up kappa?

params18 <- setResource(params18,kappa = params18@resource_params$kappa * 100)

params18 <- steady(params18)

sim18 <- project(params18, t_max = 300)

plotBiomass(sim18)
plotBiomassObservedVsModelCustom(sim18)

```

Not too bad, let's run optim on it

Optim with truncated biomass and varying rmax, erepro and dele interaction matrix


```{r}
# create set of params for the optimisation process
tic()
params_optim <- params18

vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@interaction[1,])

species = "D.ele"
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim, species = species, 
                                             dat = params_optim@species_params$biomass_observed, method ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),rep(.1,9)),
                                             upper= c(rep(15,9),rep(1,9),rep(.99,9)),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_truncatedB.RDS")
optim_result <- readRDS("params/optim_result_truncatedB.RDS")
```


```{r}
#put these new vals intospecies_params and go back to the top of this page to re-check the calibration 
species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
  interaction <- params_optim@interaction
  interaction[species,] <- optim_result$par[19:27]
  params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 2000)

plotBiomass(sim_optim)

plotBiomassObservedVsModelCustom(sim_optim)

```

RUnning sim18 through optim found a steady state but some species are pretty far from biomass data. Need to try again with lower kappa and/or adding kappa to optimisation.



```{r}

params18 <- calibrateBiomassCustom(sim17@params)

params18 <- matchBiomassCustom(params18)

sim18 <- project(params18, t_max = 500)

plotBiomassObservedVsModelCustom(sim18@params)

plotlyBiomass(sim18)
```






## Adding temperature to mizer


The temperature dependence was on:

- intake

- metabolism

- background mortality

Applied by changing:

- search volume

- intake max

- metabolism

- mu_b

Using activation and deactivation periods and a temperature reference and temperature preferred

So lots of combinations possible here. For the sake of simplicity I'm going to start with the same setup as my PhD: both activation and deactivation phases (important for heatwaves) but only on food intake (search volumme and intake max impacted)

This means I need to edit

- mizerEncounter for search_vol

- mizerFeedingLevel for intake_max

- mizerPredRate for search_vol

### Temperature functions

Code is saved in temperatureFuncs.R

Assumption: PredRate uses feeding level which uses encounter where search volume is multiplied by the scalar. The same search volume is also multiplied by the scalar in predRate. Is it ok? Guessing yes since predrate already has search volume squared in its calculation so it should be scalar squared too.


Testing the behavior of the function


```{r}

params <- newTraitParams(no_sp = 4)

params@species_params$Ea <- 0.63
params@species_params$c_a <- 0
params@species_params$Ed <- 6.3
params@species_params$c_d <- 0
params@species_params$t_d <- 25
params@species_params$t_ref <- 10

other_params(params)$temperature <- 1:30

params <- setParams(params)

params@other_params$other$temperature

temperature <- c(25)#,26)

scalar <- tempFun(params@w, temperature = temperature,
                  t_d = params@species_params$t_d,
                  t_ref = params@species_params$t_ref,
                  Ea = params@species_params$Ea,
                  c_a = params@species_params$c_a,
                  Ed = params@species_params$Ed,
                  c_d = params@species_params$c_d)
                  
scalar <- t(drop(scalar))


params@search_vol * t(scalar)



#####


w <- params@w

y = tempFun(w = w, temperature = temperature, Ea = Ea, t_ref = t_ref, Ed = Ed)

plot_dat <- data.frame("temperature" = temperature, "scalar" = y[1,])

ggplot(plot_dat) +
  geom_line(aes(x = temperature, y = scalar))

y = tempFun(w = w, temperature = 15, Ea = Ea, t_ref = t_ref, Ed = Ed)

```

Happy


test

```{r}
var <- NULL
for(t in 1:30)
{
varTemp <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]
var <- c(var,varTemp)
}

plot_dat <- data.frame("time" = 1:30, "value" = var)

var2 <- mizerEncounter(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]

ggplot(plot_dat) +
  geom_point(aes(x = time, y = value)) +
  geom_hline(yintercept = var2) # tref is 10


encounter <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = 15)

feeding_level <- mizerFeedingLevelTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       encounter = encounter, t = 15)

predRate <- mizerPredRateTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       feeding_level = feeding_level, t = 15)

```


Now writing a proper setup to enable temperature in mizer


```{r}
source("temperatureFuncs.R")
# a function that:
# adds the temperature parameters to the params object
# replace default rates in mizer with temperature mizer
# add temperature vector to param object

#' @paramTemp a dataframe with the temperature parameters to add to the mizer object 

mizerTemp <- function(temp_vec, params, paramTemp = NULL){
  
if(!is.null(paramTemp)){
params@species_params$Ea <- paramTemp$Ea
params@species_params$c_a <- paramTemp$c_a
params@species_params$Ed <- paramTemp$Ed
params@species_params$c_d <- paramTemp$c_d
params@species_params$t_d <- paramTemp$t_d
params@species_params$t_ref <- paramTemp$t_ref
}

other_params(params)$temperature <- temp_vec

params <- setRateFunction(params, "Encounter", "mizerEncounterTemp")
params <- setRateFunction(params, "FeedingLevel", "mizerFeedingLevelTemp")
params <- setRateFunction(params, "PredRate", "mizerPredRateTemp")

params <- setParams(params)
  
  return(params)
}


```


test


```{r}

params <- newTraitParams(no_sp = 4)
sim <- project(params)
plotBiomass(sim)

temperatureParams <- data.frame("Ea" = 0.63, "c_a" = 0, "Ed" = 6.3, "c_d" = 0,
                                "t_d" = 25, "t_ref" = 10)

temperature <- seq(10,30,length.out = 200)

paramsT <- mizerTemp(temp_vec = temperature, paramTemp = temperatureParams, params = params)

simT <- project(paramsT, t_max = length(temperature))

plotBiomass(simT)
```



Working :)


toothfish

Preferred temperature (Ref. 115969): 1.8 - 8.8, mean 4.5 °C (based on 1109 cells).


test

```{r}

ggplot(myvar) +
  geom_histogram(aes(x = TL)) +
  facet_wrap(~Year)



```

