---
title: toothfish project
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(mizerHowTo)
library(parallel)
library(tictoc)
load("size_spec_inputs.RData")
```

### Helpful functions


```{r}
# function for calibration of Rmax

fastOptim <- function(params)
{

# create set of params for the optimisation process
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) # variable to explore
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError,params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

params_optim@species_params$R_max <- 10^optim_result$par 

sim_optim <- project(params_optim, t_max = 2000)

return(sim_optim)
}
```


```{r}
# function running tuneParams function in a row for a quick start to a calibration
fastCalib <- function(params, match = F)
{
params <- calibrateBiomass(params) # changes kappa and rmax
if(match) params <- matchBiomasses(params) # set rmax to inf and adjust erepro
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 1000)
return(sim)
}

```


```{r}
# removing effort = 1 so always using intial effort and removing /1e6 so everything is in grams

getError2 <- function (vary, params, dat, data_type = "catch", tol = 0.1, 
    timetorun = 10) 
{
    params@species_params$R_max[] <- 10^vary[1:length(params@species_params$R_max)]
    params <- setParams(params)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    sim <- project(params, t_max = timetorun, progress_bar = F)
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}


```


```{r}

getErrorTime <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]

  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred) - log(obs))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}


```


```{r}

getErrorCustomYield <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]

  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  # biomass <- biomass[1:13,,]
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  # f_gear <- f_gear[1:13,,,,drop=F]
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred[1:13]) - log(obs[1:13]))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}


```

### Set of functions using size selectivity of the data

```{r}

# this function adds a lower boundary to selected size

plotBiomassObservedVsModelCustom <- function (object, species = NULL, ratio = FALSE, log_scale = TRUE, 
    return_data = FALSE, labels = TRUE, show_unobserved = FALSE) 
{
    if (is(object, "MizerSim")) {
        params = object@params
        n <- finalN(object)
    }
    else if (is(object, "MizerParams")) {
        params = object
        n <- initialN(params)
    }
    else {
        stop("You have not provided a valid mizerSim or mizerParams object.")
    }
    sp_params <- params@species_params
    species = valid_species_arg(object, species)
    if (length(species) == 0) 
        stop("No species selected, please fix.")
    row_select = match(species, sp_params$species)
    if (!"biomass_observed" %in% names(sp_params)) {
        stop("You have not provided values for the column 'biomass_observed' ", 
            "in the mizerParams/mizerSim object.")
    }
    else if (!is.numeric(sp_params$biomass_observed)) {
        stop("The column 'biomass_observed' in the mizerParams/mizerSim object", 
            " is not numeric, please fix.")
    }
    else {
        biomass_observed = sp_params$biomass_observed
    }
    
    cutoffLow <- sp_params$biomass_cutoffLow[row_select]
    if (is.null(cutoffLow)) {
        cutoffLow = rep(0, length(species))
    }
    else if (!is.numeric(cutoffLow)) {
        stop("params@species_params$biomass_cutoffLow is not numeric, \",\n                 \"please fix.")
    }
    cutoffLow[is.na(cutoffLow)] <- 0
    
    cutoffHigh <- sp_params$biomass_cutoffHigh[row_select]
    if (is.null(cutoffHigh)) {
        cutoffHigh = rep(0, length(species))
    }
    else if (!is.numeric(cutoffHigh)) {
        stop("params@species_params$biomass_cutoffHigh is not numeric, \",\n                 \"please fix.")
    }
    cutoffHigh[is.na(cutoffHigh)] <- 0
    
    sim_biomass = rep(0, length(species))
    for (j in 1:length(species)) {
        sim_biomass[j] = sum((n[row_select[j], ] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    dummy = data.frame(species = species, model = sim_biomass, 
        observed = biomass_observed[row_select]) %>% mutate(species = factor(species, 
        levels = species), is_observed = !is.na(observed) & observed > 
        0, observed = case_when(is_observed ~ observed, !is_observed ~ 
        model), ratio = model/observed)
    if (sum(dummy$is_observed) == 0) {
        cat(paste("There are no observed biomasses to compare to model,", 
            "only plotting model biomasses.", sep = "\n"))
    }
    if (!show_unobserved) {
        dummy <- filter(dummy, is_observed)
    }
    if (return_data == TRUE) 
        return(dummy)
    tre <- round(sum(abs(1 - dummy$ratio)), digits = 3)
    caption <- paste0("Total relative error = ", tre)
    if (any(!dummy$is_observed)) {
        caption <- paste(caption, "\n Open circles represent species without biomass observation.")
    }
    if (ratio == FALSE) {
        gg <- ggplot(data = dummy, aes(x = observed, y = model, 
            colour = species, shape = is_observed)) + geom_abline(aes(intercept = 0, 
            slope = 1), colour = "purple", linetype = "dashed", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass [g]") + 
            coord_cartesian(ylim = range(dummy$model, dummy$observed))
    }
    else {
        gg <- ggplot(data = dummy, aes(x = observed, y = ratio, 
            colour = species, shape = is_observed)) + geom_hline(aes(yintercept = 1), 
            linetype = "dashed", colour = "purple", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass / observed biomass") + 
            coord_cartesian(ylim = range(dummy$ratio))
    }
    gg <- gg + labs(x = "observed biomass [g]", caption = caption) + 
        scale_colour_manual(values = getColours(params)[dummy$species]) + 
        scale_shape_manual(values = c(`TRUE` = 19, `FALSE` = 1)) + 
        guides(shape = "none")
    if (log_scale == TRUE & ratio == FALSE) {
        gg = gg + scale_x_log10() + scale_y_log10()
    }
    if (log_scale == TRUE & ratio == TRUE) {
        gg = gg + scale_x_log10()
    }
    if (labels == TRUE) {
        gg = gg + ggrepel::geom_label_repel(aes(label = species), 
            box.padding = 0.35, point.padding = 0.5, segment.color = "grey50", 
            show.legend = FALSE, max.overlaps = Inf, seed = 42)
    }
    gg
}
```


```{r}

# adapting cutoff here too

calibrateBiomassCustom <- function (params) 
{
  if ((!("biomass_observed" %in% names(params@species_params))) || 
      all(is.na(params@species_params$biomass_observed))) {
    return(params)
  }
  no_sp <- nrow(params@species_params)
  
  cutoffLow <- params@species_params$biomass_cutoffLow
  if (is.null(cutoffLow)) 
    cutoffLow <- rep(0, no_sp)
  cutoffLow[is.na(cutoffLow)] <- 0
  
  cutoffHigh <- params@species_params$biomass_cutoffHigh
  if (is.null(cutoffHigh)) 
    cutoffHigh <- rep(0, no_sp)
  cutoffHigh[is.na(cutoffHigh)] <- 0
  
  observed <- params@species_params$biomass_observed
  observed_total <- sum(observed, na.rm = TRUE)
  sp_observed <- which(!is.na(observed))
  model_total <- 0
  for (sp_idx in sp_observed) {
    model_total <- model_total + sum((params@initial_n[sp_idx, 
    ] * params@w * params@dw)[params@w >= cutoffLow[sp_idx] & cutoffHigh[sp_idx] >= params@w])
  }
  scaleModel(params, factor = observed_total/model_total)
}


```

```{r}

# same as above

matchBiomassCustom <- function (params, species = NULL) 
{
    if (!("biomass_observed" %in% names(params@species_params))) {
        return(params)
    }
    species <- valid_species_arg(params, species = species, return.logical = TRUE) & 
        !is.na(params@species_params$biomass_observed) & params@species_params$biomass_observed > 
        0
    for (sp in (1:nrow(params@species_params))[species]) {
        cutoffLow <- params@species_params$biomass_cutoffLow[[sp]]
        if (is.null(cutoffLow) || is.na(cutoffLow)) {
            cutoffLow <- 0
        }
                    cutoffHigh <- params@species_params$biomass_cutoffHigh[[sp]]
        if (is.null(cutoffHigh) || is.na(cutoffHigh)) {
            cutoffHigh <- 0
        }
        
        total <- sum((params@initial_n[sp, ] * params@w * params@dw)[params@w >= cutoffLow & cutoffHigh >= params@w])
        factor <- params@species_params$biomass_observed[[sp]]/total
        params@initial_n[sp, ] <- params@initial_n[sp, ] * factor
    }
    setBevertonHolt(params)
}



```


```{r}
getErrorCustom <- function(vary, params, dat, tol = 0.001, 
    timetorun = 10)
{
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@species_params$interaction_resource[1:9] <- vary[19:27]

  params <- setParams(params)

  interaction <- params@interaction
  interaction[] <- matrix(vary[28:108],nrow = 9) # stop at 54 if looking only at 3 biggest species
  
  params <- setInteraction(params,interaction)

    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    sim_biomass = rep(0, length(params@species_params$species))
    
        cutoffLow <- params@species_params$biomass_cutoffLow
    if (is.null(cutoffLow)) 
        cutoffLow <- rep(0, no_sp)
    cutoffLow[is.na(cutoffLow)] <- 0
    
        cutoffHigh <- params@species_params$biomass_cutoffHigh
    if (is.null(cutoffHigh)) 
        cutoffHigh <- rep(0, no_sp)
    cutoffHigh[is.na(cutoffHigh)] <- 0
        
    for (j in 1:length(sim_biomass)) {
        sim_biomass[j] = sum((sim@n[dim(sim@n)[1],j,] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    

    pred <- log(sim_biomass)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```


For setup, run all the blocks above. Do not knit.
Latest simulation calibrated is sim18 (saved as params/params7.RDS)

### Calibration - previous steps in report.Rmd

```{r}
sim20 <- readRDS("params/params8.RDS")
# create set of params for the optimisation process
tic()
params_optim <- sim20@params

vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@species_params$interaction_resource,
        params_optim@interaction)

params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim,
                                             dat = params_optim@species_params$biomass_observed, method ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-10,9),rep(.1,9),rep(.1,81)),
                                             upper= c(rep(15,9),rep(1,9),rep(.99,9),rep(.99,81)),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_truncatedB3.RDS")
optim_result <- readRDS("params/optim_result_truncatedB2.RDS")
```


```{r}

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource <- optim_result$par[19:27]
  interaction <- params_optim@interaction
  interaction[] <- matrix(optim_result$par[28:108],nrow = 9)
  params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 2000)

plotBiomass(sim_optim)

plotBiomassObservedVsModelCustom(sim_optim)

plotDiet2(sim_optim)
plotGrowthCurves(sim_optim, species_panel = T)
```

Growth are a bit too quick... but look at that match <3 (using optim_result_truncatedB2.RDS). I did another run starting from this steady state but varying the entire interaction matrix (worked in an hour and a half on super computer) and somehow the new steady state is not as good. Don't know why but will work with params9 from now on (the perfect match)

```{r}

params_optim@initial_n <- sim_optim@n[2000,,]
sim <- project(params_optim)
saveRDS(sim, file = "params/params9.RDS")

```

Now to do the fisheries time series


```{r}

sim21 <- readRDS("params/params9.RDS")

params <- sim21@params

ll_area_yr <- read_csv("ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)
ll_km2$catchT <- ll_km2$catchT * 1e6 #in grams
ll_avg <- mean(ll_km2$catchT)
effort_avg <-mean(ll_km2$effort)

# going to use pre 2015 data only for now
# ll_km2 <- ll_km2[2:18,]

  effort <-array(ll_km2$effort, dim = c(length(ll_km2$effort),1), 
                 dimnames = list("time" = ll_km2$year, "gear" = params@gear_params$gear[1]))

# ggplot(ll_km2) +
#   geom_line(aes(x = year, y = catchT), color = "red") +
#   geom_line(aes(x = year, y = effort/1000), color = "black")




params@species_params$yield_observed <- c(ll_avg,rep(0,8))

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(7e-7,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll
initial_effort(params) <- effort_avg # somehow doesn't work with gear_params

sim22 <- project(params, t_max = 200)
# sim23 <- project(params, effort = effort)
#  
# plotBiomass(sim23)
plotFMort(sim22)

plotYieldObservedVsModel(sim22, species = "D.ele")
plotBiomass(sim22)

getYield(sim22)[200,]
sim22@params@species_params$yield_observed

```

Estimating catchability using average fisheries mortality. c = 7e-7 for close match. Going to use this starting value to do the time series


```{r}
tic()
dat <- ll_km2

  effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

# create set of params for the optimisation process
params_optim <- sim22@params
vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@gear_params$catchability[1])
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorTime,params=params_optim, dat = ll_km2, effort = effort,
                                             method   ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),1e-9),upper= c(rep(15,9),rep(1,9),1e-2),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_timeseries.RDS")
```


```{r}
optim_result <- readRDS("params/optim_result_timeseriesShort.RDS")
#put these new vals intospecies_params and go back to the top of this page to re-check the calibration 
species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1] <-optim_result$par[19]
#re-run time-varying effort model tthough time with new erepro
sim_optim <- project(params_optim, effort = effort)

plotBiomass(sim_optim)
plotFMort(sim_optim)
res <- plotYield(sim_optim,return_data = T)
res$ObservedYield <- ll_km2$catchT

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")

saveRDS(sim_optim, file = "params/params10.RDS")

```


Checking the error on the match, it's pretty big but I guess it will do

```{r}

biomass <- sweep(sim_optim@n, 3, sim_optim@params@w * sim_optim@params@dw, "*")


f_gear<-getFMortGear(params,effort)
yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                            c(1, 2, 3), sum)

yield_species <-apply(yield_species_gear, c(1, 3), sum)

yield_frame <- melt(yield_species)

obs<-dat$catchT
pred<-yield_frame$value[1:18] # only selecting D.ele for now


error <- sum((log(pred[1:13]) - log(obs[1:13]))^2,na.rm=T)

error
```


## thermizer
#### Parameters

```{r}
library(therMizer)

sim <- readRDS("params/params10.RDS")
params <- sim@params
first_year <- 1950
last_year <- 2093
times <- seq(first_year,last_year, by = 1)

# Setting up temperature range

species_params(params)$temp_min <- c(1.8,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.8,1.6)
# species_params(params)$temp_min <- c(1.4,-1.1,1.3,1.3,-0.1,-1.8,1.4,1.4,1.4) # adjusted so species survive historical temperatures

# species_params(params)$temp_max <- c(8.8,2.9,2.5,2.8,1.4,2,2.3,3.1,3.9)
species_params(params)$temp_max <- c(8.8,2.9,2.5,2.8,2.4,2.4,2.3,3.1,3.9)

```


The parameter `t_idx` will help with the simulations by providing the correct time index for the plankton and temperature forcing during the simulations.

``` {r}

# Time indexing parameter
# This will be added to t to convert the year into an index for the n_pp and ocean_temp arrays
# if (min(times) == 0) {
#   other_params(params)$t_idx = 1
# } else if (min(times) == 1) {
#   other_params(params)$t_idx = 0
# } else {
#   other_params(params)$t_idx = -(min(times) - 1)
# }
# not sure about that function, just need that:

other_params(params)$t_idx = 1

```

To scale the effect of temperature on encounter rate to a value ranging from 0 - 1, it is necessary to divide by the maximum possible value for each species.  To scale the effect of temperature on metabolism to a value ranging from 0 - 1, it is necessary to subtract the minimum vaule for each species and then divide by the range.  This requires a bit of straightforward arithmetic, and users could do this on their end if they're so inclined.  These parameters handle that math so the user doesn't have to.


```{r}
# Create parameter for scaling encounter and mortality rates
species_params(params)$encounterpred_scale <- 
    rep(NA, nrow(species_params(params)))

for (indv in seq(1:length(species_params(params)$temp_min))) {
    
    # Create a vector of all temperatures each species might encounter
    temperature <- seq(species_params(params)$temp_min[indv], 
                       species_params(params)$temp_max[indv], 
                       by = 0.1)
    
    # Find the maximum value of the unscaled effect of temperature on encounter
    # and predation rate for each species 
  species_params(params)$encounterpred_scale[indv] <- 
      max((temperature) * 
              (temperature - species_params(params)$temp_min[indv]) *
              (species_params(params)$temp_max[indv] - temperature)
          )
}

# Determine the minimum, maximum, and range of value for the effect of 
# temperature on metabolism
    
min_metab_value <- 
    (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_min)))))
max_metab_value <- 
    (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_max)))))
    
species_params(params)$metab_min <- min_metab_value
species_params(params)$metab_range <- max_metab_value - min_metab_value
```


#### Functions

Temperature will be added to the functions that determine encounter rate and energy available for growth and reproduction.

To scale encounter rate with temperature, we're essentially taking a temperature-dependent proportion of the value calculated by the mizerEncounter function.  If species are at their thermal optimum, we take the full value.  Elsewhere in their thermal range, we take a proportion that goes to zero at the limits of species' thermal tolerence.  
```{r message = FALSE, warning = FALSE}

# therMizerEncounter <- function(params, n, n_pp, n_other, t, ...) {
#   # Access the correct element
#   temp_at_t <- params@other_params$other$ocean_temp[t + params@other_params$other$t_idx,]
#   
#   # Calculate unscaled temperature effect using a generic polynomial rate equation
#   unscaled_temp_effect <- temp_at_t * (temp_at_t - params@species_params$temp_min) * (params@species_params$temp_max - temp_at_t)
#   
#   # Scale using encounter_scale parameter
#   scaled_temp_effect <- unscaled_temp_effect / params@species_params$encounter_scale
#   
#   # Set the encounter rate to zero if temperature is outside species' thermal tolerance
#   above_max <- which(temp_at_t > params@species_params$temp_max)
#   below_min <- which(temp_at_t < params@species_params$temp_min)
#   
#   if (length(above_max) > 0)
#     scaled_temp_effect[above_max] = 0
#   
#   if (length(below_min) > 0)
#     scaled_temp_effect[below_min] = 0
#   
#   # Calculate maximum possible encounter rate
#   max_encounter <- mizerEncounter(params, n = n, n_pp = n_pp, n_other = n_other, ...)
#   max_encounter <- mizerEncounter(params, n = params@initial_n, n_pp = params@other_params$other$n_pp_array[1,])
#   
#   # Apply temperature effect
#   return(max_encounter * scaled_temp_effect)
#       
# }

```

To calculate the effect of temperature on metabolim, we use an Arrhenius function to scale the cost of metabolism.  When species are at their thermal maximum, the cost of metabolism is at its maximum.  When species are at their thermal minimum, the cost of metabolism is at its minimum

```{r}
# Calculate the temperature scaling factor for the encounter rate and 
# predation rate
scaled_temp_fun <- function(params, t) {
    # Using t+1 to avoid calling ocean_temp[0,] at the first time step
# print("t in scaled_temp")
#   print(t)
#   
#   print("ocean temp")
#   print(other_params(params)$ocean_temp)
  
    temp_at_t <- other_params(params)$ocean_temp[t + params@other_params$other$t_idx,]
    
    # print("temperature in scaled_temp")
    # print(temp_at_t)
    
    # Calculate unscaled temperature effect using a generic polynomial rate equation
    unscaled_temp_effect <- 
        temp_at_t * (temp_at_t - species_params(params)$temp_min) * 
        (species_params(params)$temp_max - temp_at_t)
    
    # Scale using new parameter
    scaled_temp_effect <- 
        unscaled_temp_effect / species_params(params)$encounterpred_scale
    # print("intermediate temp effect")
    # print(scaled_temp_effect)
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    scaled_temp_effect[above_max | below_min] = 0
    
    # print("temp effect")
    # print(scaled_temp_effect)
    
    return(scaled_temp_effect)
}


therMizerEncounter <- function(params, t, ...) {
    
      # Calculate maximum possible encounter rate
      max_encounter <- mizerEncounter(params, t, ...)
      
      # Apply temperature effect
      return(max_encounter * scaled_temp_fun(params, t))
      
}

therMizerPredRate <- function(params, t, ...) {
      # Calculate maximum possible encounter rate
      max_predrate <- mizerPredRate(params, t, ...)
      # print("predrate")
      # print(max_predrate)
      # Apply temperature effect
      return(max_predrate * scaled_temp_fun(params, t))
      
}


therMizerEReproAndGrowth <- function(params, t, encounter, feeding_level, ...) {
    
  # print("t")
  # print(t)
  # print("emcounter")
  # print(encounter)
  # print("feeding")
  # print(feeding_level)
  
    # Using t+1 to avoid calling ocean_temp[0,] at the first time step
    temp_at_t <- other_params(params)$ocean_temp[t + 1,]
  
    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))
    
    # Arrhenius equation scaled to a value between 0 and 1
    temp_effect_metabolism <- 
        (unscaled_temp_effect - species_params(params)$metab_min) /
        species_params(params)$metab_range
    
    # Set temperature effect to 0 if temperatures are outside thermal 
    # tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    temp_effect_metabolism[above_max | below_min] = 0
  
  # Apply scaled Arrhenius value to metabolism
    sweep((1 - feeding_level) * encounter, 1,
          species_params(params)$alpha, "*", check.margin = FALSE) - 
        metab(params)*temp_effect_metabolism  
      
}
```

Now we'll write a function to use the CMIP6 plankton densities rather than the mizer resource dynamics.  Code for this was informed by the approach used in: <https://rpubs.com/gustav/plankton-anchovy>.  

``` {r}

# Set up new resource forcing "function" - just changes to different time slot in the n_pp_array array
plankton_forcing <- function(params, t, ...) {
  w_cut_off <- 10
  pkt <- 10^(other_params(params)$n_pp_array[t + params@other_params$other$t_idx,])/params@dw_full # converting to density
  pkt[which(as.numeric(names(pkt)) >= w_cut_off)] <- 0
  
  return(pkt)  
}

```


```{r}

# pkt_rom <- function(params, t, ...){
#   a <- rep(1,209)
#   return(a)
# }


```



Set the new rate functions and new resource

```{r}

params <- setRateFunction(params, "Encounter", "therMizerEncounter")
params <- setRateFunction(params, "PredRate", "therMizerPredRate")
params <- setRateFunction(params, "EReproAndGrowth", "therMizerEReproAndGrowth")

params <- setResource(params, resource_dynamics = "plankton_forcing")
# params <- setResource(params, resource_dynamics = "pkt_rom")

```


Editing params

```{r}

# Create parameter objects
params_IPSL_picontrol <- params
params_IPSL_ssp1rcp26 <- params
params_IPSL_ssp5rcp85 <- params

# Attach temperature
other_params(params_IPSL_picontrol)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_PIcontrol.RDS")
other_params(params_IPSL_ssp1rcp26)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_126.RDS")
other_params(params_IPSL_ssp5rcp85)$ocean_temp <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_ocean_temp_array_CCscenario_585.RDS")

# Attach plankton
other_params(params_IPSL_picontrol)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_PIcontrol.RDS")
other_params(params_IPSL_ssp1rcp26)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_126.RDS")
other_params(params_IPSL_ssp5rcp85)$n_pp_array <- readRDS("~/RstudioProjects/FISHMIP/data/processed/IPSL_n_pp_array_CCscenario_585.RDS")


```

No thermizer predrate included


Test

```{r}

sim <- project(params_IPSL_picontrol, 
               initial_n_pp = params_IPSL_picontrol@other_params$other$n_pp_array[1,],
                                      t_max = length(times))

 mizer::plotBiomass(sim)
 mizer::plotSpectra(sim)
```



Scenarios

```{r}

sim <- project(params_IPSL_ssp1rcp26, 
               initial_n_pp = params_IPSL_ssp1rcp26@other_params$other$n_pp_array[1,],
                                      t_max = length(times))

 mizer::plotBiomass(sim)
 mizer::plotSpectra(sim)
```



What does the temperature looks like?

```{r}

temp <- params_IPSL_ssp1rcp26@other_params$other$ocean_temp[,1] # all species have the same profile
temp <-temp[1:50]
plot(temp)
summary(temp)
params@species_params$temp_min
params@species_params$temp_max

plot_dat <- data.frame("Species" = rep(factor(params@species_params$species, levels = params@species_params$species),2), "temperature" = c(params@species_params$temp_min, params@species_params$temp_max))


ggplot(plot_dat) + 
  geom_line(aes(x = Species,y= temperature ), alpha = .5, size = 2)+
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = summary(temp)[2], color = "red", alpha = .5) +
  geom_hline(yintercept = summary(temp)[5], color = "red", alpha = .5) +
  geom_hline(yintercept = summary(temp)[1], color = "red", alpha = .5, linetype = "dashed") +
  geom_hline(yintercept = summary(temp)[6], color = "red", alpha = .5, linetype = "dashed") +
  scale_y_continuous(breaks = seq(0,10,.2))

```

Plot above shows the thermal range of the species with the mean temperature (between red lines) and their 1st and 3rd quantile (red dashed). Temperature data is from picon

Some species thermal range do not overlap with the current temperature. What happens in that case? (thermal range is from fishbase)

If temperature is outside species thermal range, they just die. The scalar can only scale down too...


Looking at the data

```{r}

plot_dat <- data.frame(other_params(params_IPSL_picontrol)$ocean_temp[,1], other_params(params_IPSL_ssp1rcp26)$ocean_temp[,1], other_params(params_IPSL_ssp5rcp85)$ocean_temp[,1])
colnames(plot_dat) <- c("picon", "ssp126", "ssp585")
plot_dat$Year <- as.numeric(rownames(plot_dat))

plot_dat <- reshape2::melt(plot_dat, "Year")

ggplot(plot_dat) +
  geom_line(aes(x = Year, y = value, color = variable))
```

I calibrated my species using picon but it differs from 1950 already. Need to check ssp average temperature 1950-2000 to get my temperature range

For species to survive picon I need to lower t_min but for the historical data I need to increase t_max (slighlty so they're fine with historical but not with global warming)
Funny is that the species with the widest temperature rante is toothfish and is actually perfectly fine with temperature increasing, other species are not though





