---
title: toothfish project
author: RF
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(mizerHowTo)
library(parallel)
library(tictoc)
load("size_spec_inputs.RData")
```

### Helpful functions


```{r}
# function for calibration of Rmax

fastOptim <- function(params)
{

# create set of params for the optimisation process
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) # variable to explore
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError,params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

params_optim@species_params$R_max <- 10^optim_result$par 

sim_optim <- project(params_optim, t_max = 2000)

return(sim_optim)
}
```


```{r}
# function running tuneParams function in a row for a quick start to a calibration
fastCalib <- function(params, match = F)
{
params <- calibrateBiomass(params) # changes kappa and rmax
if(match) params <- matchBiomasses(params) # set rmax to inf and adjust erepro
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 1000)
return(sim)
}

```


```{r}
# removing effort = 1 so always using intial effort and removing /1e6 so everything is in grams

getError2 <- function (vary, params, dat, data_type = "catch", tol = 0.1, 
    timetorun = 10) 
{
    params@species_params$R_max[] <- 10^vary[1:length(params@species_params$R_max)]
    params <- setParams(params)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    sim <- project(params, t_max = timetorun, progress_bar = F)
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}


```


```{r}

getErrorTime <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]

  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred) - log(obs))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}


```


```{r}

getErrorCustomYield <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]

  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  # biomass <- biomass[1:13,,]
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  # f_gear <- f_gear[1:13,,,,drop=F]
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred[1:13]) - log(obs[1:13]))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}


```

### Set of functions using size selectivity of the data

```{r}

# this function adds a lower boundary to selected size

plotBiomassObservedVsModelCustom <- function (object, species = NULL, ratio = FALSE, log_scale = TRUE, 
    return_data = FALSE, labels = TRUE, show_unobserved = FALSE) 
{
    if (is(object, "MizerSim")) {
        params = object@params
        n <- finalN(object)
    }
    else if (is(object, "MizerParams")) {
        params = object
        n <- initialN(params)
    }
    else {
        stop("You have not provided a valid mizerSim or mizerParams object.")
    }
    sp_params <- params@species_params
    species = valid_species_arg(object, species)
    if (length(species) == 0) 
        stop("No species selected, please fix.")
    row_select = match(species, sp_params$species)
    if (!"biomass_observed" %in% names(sp_params)) {
        stop("You have not provided values for the column 'biomass_observed' ", 
            "in the mizerParams/mizerSim object.")
    }
    else if (!is.numeric(sp_params$biomass_observed)) {
        stop("The column 'biomass_observed' in the mizerParams/mizerSim object", 
            " is not numeric, please fix.")
    }
    else {
        biomass_observed = sp_params$biomass_observed
    }
    
    cutoffLow <- sp_params$biomass_cutoffLow[row_select]
    if (is.null(cutoffLow)) {
        cutoffLow = rep(0, length(species))
    }
    else if (!is.numeric(cutoffLow)) {
        stop("params@species_params$biomass_cutoffLow is not numeric, \",\n                 \"please fix.")
    }
    cutoffLow[is.na(cutoffLow)] <- 0
    
    cutoffHigh <- sp_params$biomass_cutoffHigh[row_select]
    if (is.null(cutoffHigh)) {
        cutoffHigh = rep(0, length(species))
    }
    else if (!is.numeric(cutoffHigh)) {
        stop("params@species_params$biomass_cutoffHigh is not numeric, \",\n                 \"please fix.")
    }
    cutoffHigh[is.na(cutoffHigh)] <- 0
    
    sim_biomass = rep(0, length(species))
    for (j in 1:length(species)) {
        sim_biomass[j] = sum((n[row_select[j], ] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    dummy = data.frame(species = species, model = sim_biomass, 
        observed = biomass_observed[row_select]) %>% mutate(species = factor(species, 
        levels = species), is_observed = !is.na(observed) & observed > 
        0, observed = case_when(is_observed ~ observed, !is_observed ~ 
        model), ratio = model/observed)
    if (sum(dummy$is_observed) == 0) {
        cat(paste("There are no observed biomasses to compare to model,", 
            "only plotting model biomasses.", sep = "\n"))
    }
    if (!show_unobserved) {
        dummy <- filter(dummy, is_observed)
    }
    if (return_data == TRUE) 
        return(dummy)
    tre <- round(sum(abs(1 - dummy$ratio)), digits = 3)
    caption <- paste0("Total relative error = ", tre)
    if (any(!dummy$is_observed)) {
        caption <- paste(caption, "\n Open circles represent species without biomass observation.")
    }
    if (ratio == FALSE) {
        gg <- ggplot(data = dummy, aes(x = observed, y = model, 
            colour = species, shape = is_observed)) + geom_abline(aes(intercept = 0, 
            slope = 1), colour = "purple", linetype = "dashed", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass [g]") + 
            coord_cartesian(ylim = range(dummy$model, dummy$observed))
    }
    else {
        gg <- ggplot(data = dummy, aes(x = observed, y = ratio, 
            colour = species, shape = is_observed)) + geom_hline(aes(yintercept = 1), 
            linetype = "dashed", colour = "purple", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass / observed biomass") + 
            coord_cartesian(ylim = range(dummy$ratio))
    }
    gg <- gg + labs(x = "observed biomass [g]", caption = caption) + 
        scale_colour_manual(values = getColours(params)[dummy$species]) + 
        scale_shape_manual(values = c(`TRUE` = 19, `FALSE` = 1)) + 
        guides(shape = "none")
    if (log_scale == TRUE & ratio == FALSE) {
        gg = gg + scale_x_log10() + scale_y_log10()
    }
    if (log_scale == TRUE & ratio == TRUE) {
        gg = gg + scale_x_log10()
    }
    if (labels == TRUE) {
        gg = gg + ggrepel::geom_label_repel(aes(label = species), 
            box.padding = 0.35, point.padding = 0.5, segment.color = "grey50", 
            show.legend = FALSE, max.overlaps = Inf, seed = 42)
    }
    gg
}
```


```{r}

# adapting cutoff here too

calibrateBiomassCustom <- function (params) 
{
  if ((!("biomass_observed" %in% names(params@species_params))) || 
      all(is.na(params@species_params$biomass_observed))) {
    return(params)
  }
  no_sp <- nrow(params@species_params)
  
  cutoffLow <- params@species_params$biomass_cutoffLow
  if (is.null(cutoffLow)) 
    cutoffLow <- rep(0, no_sp)
  cutoffLow[is.na(cutoffLow)] <- 0
  
  cutoffHigh <- params@species_params$biomass_cutoffHigh
  if (is.null(cutoffHigh)) 
    cutoffHigh <- rep(0, no_sp)
  cutoffHigh[is.na(cutoffHigh)] <- 0
  
  observed <- params@species_params$biomass_observed
  observed_total <- sum(observed, na.rm = TRUE)
  sp_observed <- which(!is.na(observed))
  model_total <- 0
  for (sp_idx in sp_observed) {
    model_total <- model_total + sum((params@initial_n[sp_idx, 
    ] * params@w * params@dw)[params@w >= cutoffLow[sp_idx] & cutoffHigh[sp_idx] >= params@w])
  }
  scaleModel(params, factor = observed_total/model_total)
}


```

```{r}

# same as above

matchBiomassCustom <- function (params, species = NULL) 
{
    if (!("biomass_observed" %in% names(params@species_params))) {
        return(params)
    }
    species <- valid_species_arg(params, species = species, return.logical = TRUE) & 
        !is.na(params@species_params$biomass_observed) & params@species_params$biomass_observed > 
        0
    for (sp in (1:nrow(params@species_params))[species]) {
        cutoffLow <- params@species_params$biomass_cutoffLow[[sp]]
        if (is.null(cutoffLow) || is.na(cutoffLow)) {
            cutoffLow <- 0
        }
                    cutoffHigh <- params@species_params$biomass_cutoffHigh[[sp]]
        if (is.null(cutoffHigh) || is.na(cutoffHigh)) {
            cutoffHigh <- 0
        }
        
        total <- sum((params@initial_n[sp, ] * params@w * params@dw)[params@w >= cutoffLow & cutoffHigh >= params@w])
        factor <- params@species_params$biomass_observed[[sp]]/total
        params@initial_n[sp, ] <- params@initial_n[sp, ] * factor
    }
    setBevertonHolt(params)
}



```


```{r}
getErrorCustom <- function(vary, params, dat, tol = 0.001, 
    timetorun = 10)
{
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@species_params$interaction_resource[1:9] <- vary[19:27]

  params <- setParams(params)

  interaction <- params@interaction
  interaction[] <- matrix(vary[28:108],nrow = 9) # stop at 54 if looking only at 3 biggest species
  
  params <- setInteraction(params,interaction)

    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    sim_biomass = rep(0, length(params@species_params$species))
    
        cutoffLow <- params@species_params$biomass_cutoffLow
    if (is.null(cutoffLow)) 
        cutoffLow <- rep(0, no_sp)
    cutoffLow[is.na(cutoffLow)] <- 0
    
        cutoffHigh <- params@species_params$biomass_cutoffHigh
    if (is.null(cutoffHigh)) 
        cutoffHigh <- rep(0, no_sp)
    cutoffHigh[is.na(cutoffHigh)] <- 0
        
    for (j in 1:length(sim_biomass)) {
        sim_biomass[j] = sum((sim@n[dim(sim@n)[1],j,] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    

    pred <- log(sim_biomass)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```


For setup, run all the blocks above. Do not knit.
Latest simulation calibrated is sim18 (saved as params/params7.RDS)

### Calibration - previous steps in report.Rmd

```{r}
sim20 <- readRDS("params/params8.RDS")
# create set of params for the optimisation process
tic()
params_optim <- sim20@params

vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@species_params$interaction_resource,
        params_optim@interaction)

params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorCustom,params=params_optim,
                                             dat = params_optim@species_params$biomass_observed, method ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-10,9),rep(.1,9),rep(.1,81)),
                                             upper= c(rep(15,9),rep(1,9),rep(.99,9),rep(.99,81)),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_truncatedB3.RDS")
optim_result <- readRDS("params/optim_result_truncatedB2.RDS")
```


```{r}

species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
species_params(params_optim)$interaction_resource <- optim_result$par[19:27]
  interaction <- params_optim@interaction
  interaction[] <- matrix(optim_result$par[28:108],nrow = 9)
  params_optim <- setInteraction(params_optim,interaction)

sim_optim <- project(params_optim, t_max = 2000)

plotBiomass(sim_optim)

plotBiomassObservedVsModelCustom(sim_optim)

plotDiet2(sim_optim)
plotGrowthCurves(sim_optim, species_panel = T)
```

Growth are a bit too quick... but look at that match <3 (using optim_result_truncatedB2.RDS). I did another run starting from this steady state but varying the entire interaction matrix (worked in an hour and a half on super computer) and somehow the new steady state is not as good. Don't know why but will work with params9 from now on (the perfect match)

```{r}

params_optim@initial_n <- sim_optim@n[2000,,]
sim <- project(params_optim)
saveRDS(sim, file = "params/params9.RDS")

```

Now to do the fisheries time series


```{r}

sim21 <- readRDS("params/params9.RDS")

params <- sim21@params

ll_area_yr <- read_csv("ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)
ll_km2$catchT <- ll_km2$catchT * 1e6 #in grams
ll_avg <- mean(ll_km2$catchT)
effort_avg <-mean(ll_km2$effort)

# going to use pre 2015 data only for now
# ll_km2 <- ll_km2[2:18,]

  effort <-array(ll_km2$effort, dim = c(length(ll_km2$effort),1), 
                 dimnames = list("time" = ll_km2$year, "gear" = params@gear_params$gear[1]))

# ggplot(ll_km2) +
#   geom_line(aes(x = year, y = catchT), color = "red") +
#   geom_line(aes(x = year, y = effort/1000), color = "black")




params@species_params$yield_observed <- c(ll_avg,rep(0,8))

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(7e-7,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll
initial_effort(params) <- effort_avg # somehow doesn't work with gear_params

sim22 <- project(params, t_max = 200)
# sim23 <- project(params, effort = effort)
#  
# plotBiomass(sim23)
plotFMort(sim22)

plotYieldObservedVsModel(sim22, species = "D.ele")
plotBiomass(sim22)

getYield(sim22)[200,]
sim22@params@species_params$yield_observed

```

Estimating catchability using average fisheries mortality. c = 7e-7 for close match. Going to use this starting value to do the time series


```{r}
tic()
dat <- ll_km2

  effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

# create set of params for the optimisation process
params_optim <- sim22@params
vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@gear_params$catchability[1])
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorTime,params=params_optim, dat = ll_km2, effort = effort,
                                             method   ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),1e-9),upper= c(rep(15,9),rep(1,9),1e-2),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_timeseries.RDS")
```


```{r}
optim_result <- readRDS("params/optim_result_timeseriesShort.RDS")
#put these new vals intospecies_params and go back to the top of this page to re-check the calibration 
species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1] <-optim_result$par[19]
#re-run time-varying effort model tthough time with new erepro
sim_optim <- project(params_optim, effort = effort)

plotBiomass(sim_optim)
plotFMort(sim_optim)
res <- plotYield(sim_optim,return_data = T)
res$ObservedYield <- ll_km2$catchT

ggplot(res, aes(x = Year)) +
  geom_line(aes(y = Yield)) +
  geom_line(aes(y = ObservedYield), color = "red")

saveRDS(sim_optim, file = "params/params10.RDS")

```


Checking the error on the match, it's pretty big but I guess it will do

```{r}

biomass <- sweep(sim_optim@n, 3, sim_optim@params@w * sim_optim@params@dw, "*")


f_gear<-getFMortGear(params,effort)
yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                            c(1, 2, 3), sum)

yield_species <-apply(yield_species_gear, c(1, 3), sum)

yield_frame <- melt(yield_species)

obs<-dat$catchT
pred<-yield_frame$value[1:18] # only selecting D.ele for now


error <- sum((log(pred[1:13]) - log(obs[1:13]))^2,na.rm=T)

error
```


## Adding the temperature using therMizer

Functions to load
```{r}
library(therMizer)
# Calculate the temperature scaling factor for the encounter rate, 
# predation mortality rate and resource mortality rate
scaled_temp_effect <- function(t) {
  # Using t+1 to avoid calling ocean_temp[0,] at the first time step
  temp_at_t <- other_params(params)$ocean_temp[t + 1,]
  
  # Calculate unscaled temperature effect using a generic polynomial rate equation
  unscaled_temp_effect <- 
    temp_at_t * (temp_at_t - species_params(params)$temp_min) * 
    (species_params(params)$temp_max - temp_at_t)
  
  # Scale using new parameter
  scaled_temp_effect <- unscaled_temp_effect / species_params(params)$encounterpred_scale
  
  # Set temperature effect to 0 if temperatures are outside thermal tolerance limits
  above_max <- temp_at_t > species_params(params)$temp_max
  below_min <- temp_at_t < species_params(params)$temp_min
  
  scaled_temp_effect[above_max | below_min] = 0
  
  scaled_temp_effect
}
therMizerEncounter <- function(params, t, ...) {
  
  # Calculate maximum possible encounter rate
  max_encounter <- mizerEncounter(params, t, ...)
  
  # Apply temperature effect
  # return(sweep(max_encounter, 1, scaled_temp_effect, '*', check.margin = FALSE))
  return(max_encounter * scaled_temp_effect(t))
  
}
therMizerPredRate <- function(params, t, ...) {
  # Calculate maximum possible encounter rate
  max_predrate <- mizerPredRate(params, t, ...)
  
  # Apply temperature effect
  # return(sweep(max_predrate, 1, scaled_temp_effect, '*', check.margin = FALSE))
  return(max_predrate * scaled_temp_effect(t))
  
}

therMizerEReproAndGrowth <- function(params, t, encounter, feeding_level, ...) {
    
    # Using t+1 to avoid calling ocean_temp[0,] at the first time step
    temp_at_t <- other_params(params)$ocean_temp[t + 1,]
  
    # Arrhenius equation
    unscaled_temp_effect <- (exp(25.22 - (0.63/((8.62e-5)*(273 + temp_at_t)))))
    
    # Arrhenius equation scaled to a value between 0 and 1
        temp_effect_metabolism <- (unscaled_temp_effect - species_params(params)$metab_min) / species_params(params)$metab_range
        
        # Set temperature effect to 0 if temperatures are outside thermal tolerance limits
    above_max <- temp_at_t > species_params(params)$temp_max
    below_min <- temp_at_t < species_params(params)$temp_min
    
    temp_effect_metabolism[above_max | below_min] = 0
  
        # Apply scaled Arrhenius value to metabolism
    sweep((1 - feeding_level) * encounter, 1,
               species_params(params)$alpha, "*", check.margin = FALSE) - 
      metab(params)*temp_effect_metabolism  
      
}
```


Editing params

```{r}

sim <- readRDS("params/params10.RDS")
params <- sim@params

species_params(params)$temp_min <- rep(4, 9)
species_params(params)$temp_max <- rep(10, 9)

# Create parameter for scaling encounter and mortality rates
species_params(params)$encounterpred_scale <- rep(NA, length(species_params(params)$temp_min))
for (indv in seq(1:length(species_params(params)$temp_min))) {
  
  # Create a vector of all temperatures each species might encounter
  temperature <- seq(species_params(params)$temp_min[indv], species_params(params)$temp_max[indv], by = 0.1)
  
  # Find the maximum value of the unscaled effect of temperature on encounter and predation rate for each species 
  species_params(params)$encounterpred_scale[indv] <- max((temperature) * (temperature - species_params(params)$temp_min[indv]) * (species_params(params)$temp_max[indv] - temperature))
  
}
# Determine the minimum, maximum, and range of value for the effect of temperature on metabolism

min_metab_value <- (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_min)))))
max_metab_value <- (exp(25.22 - (0.63/((8.62e-5)*(273 + species_params(params)$temp_max)))))

species_params(params)$metab_min <- min_metab_value
species_params(params)$metab_range <- max_metab_value - min_metab_value


# Create temperature array and fill it
times <- 0:500
species <- species_params(params)$species
ocean_temp_array <- array(NA, dim = c(length(times), length(species)), dimnames = list(time = times, sp = species))
temp_inc <- 0
for (i in (times+1)) {
  ocean_temp_array[i,] <- rep(1 + temp_inc, length(species))
  temp_inc <- temp_inc + 0.01
}
other_params(params)$ocean_temp <- ocean_temp_array


params <- setRateFunction(params, "Encounter", "therMizerEncounter")
params <- setRateFunction(params, "PredRate", "therMizerPredRate")
params <- setRateFunction(params, "EReproAndGrowth", "therMizerEReproAndGrowth")

```

Test

```{r}

sim2 <- project(params, t_max = 500, effort = 0)

mizer::plotBiomass(sim2)
```

Cool it works, now I need the temperature time series to slot into ocean_temp_array



## Adding temperature to mizer


The temperature dependence was on:

- intake

- metabolism

- background mortality

Applied by changing:

- search volume

- intake max

- metabolism

- mu_b

Using activation and deactivation periods and a temperature reference and temperature preferred

So lots of combinations possible here. For the sake of simplicity I'm going to start with the same setup as my PhD: both activation and deactivation phases (important for heatwaves) but only on food intake (search volumme and intake max impacted)

This means I need to edit

- mizerEncounter for search_vol

- mizerFeedingLevel for intake_max

- mizerPredRate for search_vol

### Temperature functions

Code is saved in temperatureFuncs.R

Assumption: PredRate uses feeding level which uses encounter where search volume is multiplied by the scalar. The same search volume is also multiplied by the scalar in predRate. Is it ok? Guessing yes since predrate already has search volume squared in its calculation so it should be scalar squared too.


Testing the behavior of the function


```{r}

params <- newTraitParams(no_sp = 4)

params@species_params$Ea <- 0.63
params@species_params$c_a <- 0
params@species_params$Ed <- 6.3
params@species_params$c_d <- 0
params@species_params$t_d <- 25
params@species_params$t_ref <- 10

other_params(params)$temperature <- 1:30

params <- setParams(params)

params@other_params$other$temperature

temperature <- c(25)#,26)

scalar <- tempFun(params@w, temperature = temperature,
                  t_d = params@species_params$t_d,
                  t_ref = params@species_params$t_ref,
                  Ea = params@species_params$Ea,
                  c_a = params@species_params$c_a,
                  Ed = params@species_params$Ed,
                  c_d = params@species_params$c_d)
                  
scalar <- t(drop(scalar))


params@search_vol * t(scalar)



#####


w <- params@w

y = tempFun(w = w, temperature = temperature, Ea = Ea, t_ref = t_ref, Ed = Ed)

plot_dat <- data.frame("temperature" = temperature, "scalar" = y[1,])

ggplot(plot_dat) +
  geom_line(aes(x = temperature, y = scalar))

y = tempFun(w = w, temperature = 15, Ea = Ea, t_ref = t_ref, Ed = Ed)

```

Happy


test

```{r}
var <- NULL
for(t in 1:30)
{
varTemp <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]
var <- c(var,varTemp)
}

plot_dat <- data.frame("time" = 1:30, "value" = var)

var2 <- mizerEncounter(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]

ggplot(plot_dat) +
  geom_point(aes(x = time, y = value)) +
  geom_hline(yintercept = var2) # tref is 10


encounter <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = 15)

feeding_level <- mizerFeedingLevelTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       encounter = encounter, t = 15)

predRate <- mizerPredRateTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       feeding_level = feeding_level, t = 15)

```


Now writing a proper setup to enable temperature in mizer


```{r}
source("temperatureFuncs.R")
# a function that:
# adds the temperature parameters to the params object
# replace default rates in mizer with temperature mizer
# add temperature vector to param object

#' @paramTemp a dataframe with the temperature parameters to add to the mizer object 

mizerTemp <- function(temp_vec, params, paramTemp = NULL){
  
if(!is.null(paramTemp)){
params@species_params$Ea <- paramTemp$Ea
params@species_params$c_a <- paramTemp$c_a
params@species_params$Ed <- paramTemp$Ed
params@species_params$c_d <- paramTemp$c_d
params@species_params$t_d <- paramTemp$t_d
params@species_params$t_ref <- paramTemp$t_ref
}

other_params(params)$temperature <- temp_vec

params <- setRateFunction(params, "Encounter", "mizerEncounterTemp")
params <- setRateFunction(params, "FeedingLevel", "mizerFeedingLevelTemp")
params <- setRateFunction(params, "PredRate", "mizerPredRateTemp")

params <- setParams(params)
  
  return(params)
}


```


test


```{r}

params <- newTraitParams(no_sp = 4)
sim <- project(params)
plotBiomass(sim)

temperatureParams <- data.frame("Ea" = 0.63, "c_a" = 0, "Ed" = 6.3, "c_d" = 0,
                                "t_d" = 25, "t_ref" = 10)

temperature <- seq(10,30,length.out = 200)

paramsT <- mizerTemp(temp_vec = temperature, paramTemp = temperatureParams, params = params)

simT <- project(paramsT, t_max = length(temperature))

plotBiomass(simT)
```



Working :)


toothfish

Preferred temperature (Ref. 115969): 1.8 - 8.8, mean 4.5 °C (based on 1109 cells).


test

```{r}

ggplot(myvar) +
  geom_histogram(aes(x = TL)) +
  facet_wrap(~Year)



```

