---
title: temperature to make as package
author: RF



## Adding temperature to mizer


The temperature dependence was on:

- intake

- metabolism

- background mortality

Applied by changing:

- search volume

- intake max

- metabolism

- mu_b

Using activation and deactivation periods and a temperature reference and temperature preferred

So lots of combinations possible here. For the sake of simplicity I'm going to start with the same setup as my PhD: both activation and deactivation phases (important for heatwaves) but only on food intake (search volumme and intake max impacted)

This means I need to edit

- mizerEncounter for search_vol

- mizerFeedingLevel for intake_max

- mizerPredRate for search_vol

### Temperature functions

Code is saved in temperatureFuncs.R

Assumption: PredRate uses feeding level which uses encounter where search volume is multiplied by the scalar. The same search volume is also multiplied by the scalar in predRate. Is it ok? Guessing yes since predrate already has search volume squared in its calculation so it should be scalar squared too.


Testing the behavior of the function


```{r}

params <- newTraitParams(no_sp = 4)

params@species_params$Ea <- 0.63
params@species_params$c_a <- 0
params@species_params$Ed <- 6.3
params@species_params$c_d <- 0
params@species_params$t_d <- 25
params@species_params$t_ref <- 10

other_params(params)$temperature <- 1:30

params <- setParams(params)

params@other_params$other$temperature

temperature <- c(25)#,26)

scalar <- tempFun(params@w, temperature = temperature,
                  t_d = params@species_params$t_d,
                  t_ref = params@species_params$t_ref,
                  Ea = params@species_params$Ea,
                  c_a = params@species_params$c_a,
                  Ed = params@species_params$Ed,
                  c_d = params@species_params$c_d)
                  
scalar <- t(drop(scalar))


params@search_vol * t(scalar)



#####


w <- params@w

y = tempFun(w = w, temperature = temperature, Ea = Ea, t_ref = t_ref, Ed = Ed)

plot_dat <- data.frame("temperature" = temperature, "scalar" = y[1,])

ggplot(plot_dat) +
  geom_line(aes(x = temperature, y = scalar))

y = tempFun(w = w, temperature = 15, Ea = Ea, t_ref = t_ref, Ed = Ed)

```

Happy


test

```{r}
var <- NULL
for(t in 1:30)
{
varTemp <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]
var <- c(var,varTemp)
}

plot_dat <- data.frame("time" = 1:30, "value" = var)

var2 <- mizerEncounter(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]

ggplot(plot_dat) +
  geom_point(aes(x = time, y = value)) +
  geom_hline(yintercept = var2) # tref is 10


encounter <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = 15)

feeding_level <- mizerFeedingLevelTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       encounter = encounter, t = 15)

predRate <- mizerPredRateTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       feeding_level = feeding_level, t = 15)

```


Now writing a proper setup to enable temperature in mizer


```{r}
source("temperatureFuncs.R")
# a function that:
# adds the temperature parameters to the params object
# replace default rates in mizer with temperature mizer
# add temperature vector to param object

#' @paramTemp a dataframe with the temperature parameters to add to the mizer object 

mizerTemp <- function(temp_vec, params, paramTemp = NULL){
  
if(!is.null(paramTemp)){
params@species_params$Ea <- paramTemp$Ea
params@species_params$c_a <- paramTemp$c_a
params@species_params$Ed <- paramTemp$Ed
params@species_params$c_d <- paramTemp$c_d
params@species_params$t_d <- paramTemp$t_d
params@species_params$t_ref <- paramTemp$t_ref
}

other_params(params)$temperature <- temp_vec

params <- setRateFunction(params, "Encounter", "mizerEncounterTemp")
params <- setRateFunction(params, "FeedingLevel", "mizerFeedingLevelTemp")
params <- setRateFunction(params, "PredRate", "mizerPredRateTemp")

params <- setParams(params)
  
  return(params)
}


```


test


```{r}

params <- newTraitParams(no_sp = 4)
sim <- project(params)
plotBiomass(sim)

temperatureParams <- data.frame("Ea" = 0.63, "c_a" = 0, "Ed" = 6.3, "c_d" = 0,
                                "t_d" = 25, "t_ref" = 10)

temperature <- seq(10,30,length.out = 200)

paramsT <- mizerTemp(temp_vec = temperature, paramTemp = temperatureParams, params = params)

simT <- project(paramsT, t_max = length(temperature))

plotBiomass(simT)
```



Working :)


toothfish

Preferred temperature (Ref. 115969): 1.8 - 8.8, mean 4.5 °C (based on 1109 cells).


test

```{r}

ggplot(myvar) +
  geom_histogram(aes(x = TL)) +
  facet_wrap(~Year)



```

