---
title: toothfish project
author: RF & JB
output:
  html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizer)
library(mizerExperimental)
library(mizerHowTo)
library(parallel)
library(tictoc)
load("size_spec_inputs.RData")
```

### Helpful functions


Explore the model output of calibration only - see report.Rmd for all of Romain's calibration steps...


Latest simulation calibrated is sim16 (saved as params/params6.RDS)

### Calibration notes

The following param object is a Mizer params object assembled from different dataset from the toothfish team. So it's either empirical data or default mizer value. The interaction matrix is approximated from a cooccurence matrix of the survey, however the values are rather low so they've been multiplied by 2.





```{r}
sim16 <- readRDS("params/params6.RDS")
sim17 <- project(sim16@params, t_max = 500, effort = 1)
plotBiomass(sim17)
sim17@params@species_params$biomass_cutoff <- sim17@params@species_params$w_mat25
plotBiomassObservedVsModel(sim17)
 # need to redo biomass matching with min and max biomass - here used w_mat25 bit 
plotGrowthCurves(sim17,species_panel = T)
# shows general growth looks OK

plotDiet2(sim17,species="D.ele")
# can see toothfish change diets with growth - good - eat more fish and some cannibalism. Can llok at these for others species but showing toothfish for the stake holder talk


plotYieldVsF(sim17@params,species="D.ele",no_steps=50,F_max=1.2)
# Fmsy around 0.1 for toothfish


getYieldVsF(sim17@params,species="D.ele",no_steps=50,F_max=1.2)

# shows that in current model set up toothfish very sensitive to fishing

# suggests stock recruit relationship parameters needs looking at in more detail ( working on with more detailed calibration)

#for example, increasing Rmax or erepro makes less sensitive...but affects coexistence...

sens_test_erepro<-sim17@params
sens_test_erepro@species_params$erepro[1]<-1

sens_test1<-getYieldVsF(sens_test_erepro,species="D.ele",no_steps=50,F_max=1.2)

plotYieldVsF(sens_test_erepro,species="D.ele",no_steps=50,F_max=1.2)

sens_test_gear<-sim17@params
sens_test_gear@gear_params$knife_edge_size<-sim17@params@species_params$w_mat[1]*2

plotYieldVsF(sens_test_gear,species="D.ele",no_steps=50,F_max=1.2)


sens_test_Rmax<-sim17@params
sens_test_Rmax@species_params$R_max[1]<-sens_test_Rmax@species_params$R_max[1]*10
  
getYieldVsF(sens_test_Rmax,species="D.ele",no_steps=50,F_max=1.2)

plotYieldVsF(sens_test_Rmax,species="D.ele",no_steps=50,F_max=1.2)



```


```{r}

params<-sim17@params
params2 <- setBevertonHolt(params, reproduction_level = 0.9)

plotBevertonHolt2 <- function(params, params2, species) {
  select <- species_params(params)$species == species
  erepro <- species_params(params)$erepro[select]
  w0 <- params@w[params@w_min_idx[select]]
  E_R_ss <- getRDI(params)[select] / erepro * 2 * w0
  R_dd_ss <- getRDD(params)[select]
  E_R <- seq(0, 2 * E_R_ss, length.out = 50)
  
  R_max  <- species_params(params)$R_max[select]
  R_di = erepro * E_R / 2 / w0
  R_dd <- R_di / (1 + R_di / R_max)
  df <- melt(data.frame(E_R, R_dd, R_di, R_max), id.vars = "E_R")
  df$Model <- "Model 1"
  
  erepro <- species_params(params2)$erepro[select]
  R_max  <- species_params(params2)$R_max[select]
  R_di = erepro * E_R / 2 / w0
  R_dd <- R_di / (1 + R_di / R_max)
  df2 <- melt(data.frame(E_R, R_dd, R_di, R_max), id.vars = "E_R")
  df2$Model <- "Model 2"
  
  ggplot(rbind(df, df2)) +
    geom_line(aes(x = E_R, y = value, linetype = variable,
                  colour = Model, size = Model)) +
    geom_point(aes(x = E_R_ss, y = R_dd_ss), size = 3, color = "red") +
    ylim(NA, 1.1 * R_max) +
    ylab("Reproduction rate [eggs/year]") +
    xlab("Energy invested [g/year]") +
    labs(linetype = "", size = "R_max", colour = "R_max") +
    scale_size_manual(values = c(0.5, 1)) +
    scale_colour_manual(values = c("blue", "black")) +
    scale_linetype_manual(values = c("solid", "dashed", "dotted"))
}

plotBevertonHolt2(params, sens_test_Rmax, "D.ele")

plotYieldVsF(sens_test_Rmax,species="D.ele",no_steps=50,F_max=1.2)

```



```{r}
plotFMort(sim17)
plotPredMort(sim17)
plotSpectra(sim17)
```

Toothfish grow big and has a slow growth and yet they fish just small young individuals? WTF

Issue: how can the slow growth counterbalance the fisheries mortality?

```{r}

plot_dat <- plotEnergyBudget(sim16, species = "D.ele", return_data = T)
params <- sim16@params
 params <- setColours(params, list(Growth = "#F8766D", 
        Income = "#7CAE00", `Metabolic loss` = "#00BFCFC4", 
        Reproduction = "#C77CFF"))
xtrans <- "log10"
 
pl <- plotDataFrame(plot_dat[[1]], params, style = "area", 
        xlab = "Size [g]", ylab = "Rate [g/year]", 
        xtrans = xtrans, wrap_var = "Species", wrap_scale = "free")

    pl <- pl + geom_vline(data = plot_dat[[2]], aes(xintercept = w_mat, 
        group = Species), linetype = "dotted") + 
      geom_vline(data = plot_dat[[2]], 
        aes(xintercept = w_inf, group = Species), linetype = "dotted") + 
        geom_text(data = plot_dat[[2]], aes(x = w_mat, y = y_coord * 
            0.2, label = "\nMaturity"), angle = 90) + 
      geom_text(data = plot_dat[[2]], 
        aes(x = w_inf, y = y_coord * 0.2, label = "\nMaximum"), 
        angle = 90)+
      geom_vline(xintercept = 2722, color = "red")


```


Above graph shows that in this fisheries conditions, toothfish cannot reproduce. But the modelled growth is already slightly faster than the vonB one...


In the meantime let's recreate the longlines at a km2 scale

```{r}

ll_area_yr <- read_csv("ll_area_yr.csv")

ll_trends <- cbind(ll_trends,ll_area_yr[-dim(ll_area_yr)[1],3])

ll_km2 <- data.frame("year" = ll_trends$Year, "effort" = ll_trends$Effort/ll_trends$area_km,
                     "catchT" = ll_trends$CatchT / ll_trends$area_km)

ll_avg <- mean(ll_km2$catchT) * 1e6 # in grams
```

what does it look like?


```{r}

ggplot(ll_km2) +
  geom_line(aes(x = year, y = catchT), color = "red") +
  geom_line(aes(x = year, y = effort/1000), color = "black")

```


Redoing the fisheries and matching to time series


```{r}

params <- sim16@params

params@species_params$yield_observed <- c(ll_avg,rep(0,8))

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(0.00001,rep(0,8)),
               initial_effort = 1)

gear_params(params) <- gear_ll
initial_effort(params) <- 1 # somehow doesn't work with gear_params

sim16 <- project(params, t_max = 500)
plotBiomass(sim16)
plotFMort(sim16)


```



```{r}

params17 <- sim16@params
params17@initial_n <- sim16@n[dim(sim16@n)[1],,]
params17@initial_n_pp <- sim16@n_pp[dim(sim16@n_pp)[1],]

params17 <- setParams(params17)

sim17 <- project(params17)

plotBiomass(sim17)



params <- sim17@params

gear_ll<-data.frame(species = params@species_params$species,
               gear = "longline",
               sel_func = "knife_edge",
               knife_edge_size = 2722 ,
               catchability = c(1e-5,rep(0,8)),
               initial_effort = mean(ll_km2$effort))

gear_params(params) <- gear_ll

sim18 <- project(params, effort = params@gear_params$initial_effort[1], t_max = 1000)

plotBiomass(sim18)

plotFMort(sim18)

gear_params(sim18@params)

plotYield(sim18)
plotBiomass(sim18)

final_params<-sim18@params
initialN(final_params)<-sim18@n[1000,,]

sim<-project(final_params,effort = params@gear_params$initial_effort[1])
plotBiomass(sim)

plotYieldObservedVsModel(sim)
params@species_params



sim<-project(final_params,effort = params@gear_params$initial_effort[1])

effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

sim_time <- project(final_params, effort = effort)

plotBiomass(sim_time,species = "D.ele")
plotYield(sim_time,species = "D.ele")

modelledY<-getYield(sim_time)[,1]


plotlyGrowthCurves(sim)


plotGrowthCurves(sim, species_panel = TRUE)


plotYieldObservedVsModel(sim_time)
```




if does not work do the optim on time averaged first


At least it works, nevertheless the low catchability

Let's run it through optim varying erepro, rmax and catchability


```{r}
tic()
dat <- ll_km2

  effort <-array(dat$effort, dim = c(length(dat$effort),1), 
                 dimnames = list("time" = dat$year, "gear" = params@gear_params$gear[1]))

# create set of params for the optimisation process
params_optim <- final_params
vary<-c(log10(params_optim@species_params$R_max),
        params_optim@species_params$erepro,
        params_optim@gear_params$catchability[1])
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getErrorTime,params=params_optim, dat = ll_km2, effort = effort,
                                             method   ="L-BFGS-B", 
                                             lower=c(rep(3,9),rep(1e-7,9),0.000001),upper= c(rep(15,9),rep(1,9),1),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)

toc()
saveRDS(optim_result, file="params/optim_result_timeseries.RDS")
optim_result <- readRDS("params/optim_result_timeseries.RDS")
#put these new vals intospecies_params and go back to the top of this page to re-check the calibration 
species_params(params_optim)$R_max<-10^optim_result$par[1:9]
species_params(params_optim)$erepro<-optim_result$par[10:18]
gear_params(params_optim)$catchability[1] <-optim_result$par[19]
#re-run time-varying effort model tthough time with new erepro
sim_optim <- project(params_optim, effort = effort)




```


```{r}
plotBiomass(sim_time)
plotFMort(sim_time)
plotYield(sim_time)
```

This is not working




## Adding temperature to mizer


The temperature dependence was on:

- intake

- metabolism

- background mortality

Applied by changing:

- search volume

- intake max

- metabolism

- mu_b

Using activation and deactivation periods and a temperature reference and temperature preferred

So lots of combinations possible here. For the sake of simplicity I'm going to start with the same setup as my PhD: both activation and deactivation phases (important for heatwaves) but only on food intake (search volumme and intake max impacted)

This means I need to edit

- mizerEncounter for search_vol

- mizerFeedingLevel for intake_max

- mizerPredRate for search_vol

### Temperature functions

Code is saved in temperatureFuncs.R

Assumption: PredRate uses feeding level which uses encounter where search volume is multiplied by the scalar. The same search volume is also multiplied by the scalar in predRate. Is it ok? Guessing yes since predrate already has search volume squared in its calculation so it should be scalar squared too.


Testing the behavior of the function


```{r}

params <- newTraitParams(no_sp = 4)

params@species_params$Ea <- 0.63
params@species_params$c_a <- 0
params@species_params$Ed <- 6.3
params@species_params$c_d <- 0
params@species_params$t_d <- 25
params@species_params$t_ref <- 10

other_params(params)$temperature <- 1:30

params <- setParams(params)

params@other_params$other$temperature

temperature <- c(25)#,26)

scalar <- tempFun(params@w, temperature = temperature,
                  t_d = params@species_params$t_d,
                  t_ref = params@species_params$t_ref,
                  Ea = params@species_params$Ea,
                  c_a = params@species_params$c_a,
                  Ed = params@species_params$Ed,
                  c_d = params@species_params$c_d)
                  
scalar <- t(drop(scalar))


params@search_vol * t(scalar)



#####


w <- params@w

y = tempFun(w = w, temperature = temperature, Ea = Ea, t_ref = t_ref, Ed = Ed)

plot_dat <- data.frame("temperature" = temperature, "scalar" = y[1,])

ggplot(plot_dat) +
  geom_line(aes(x = temperature, y = scalar))

y = tempFun(w = w, temperature = 15, Ea = Ea, t_ref = t_ref, Ed = Ed)

```

Happy


test

```{r}
var <- NULL
for(t in 1:30)
{
varTemp <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]
var <- c(var,varTemp)
}

plot_dat <- data.frame("time" = 1:30, "value" = var)

var2 <- mizerEncounter(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = t)[1,50]

ggplot(plot_dat) +
  geom_point(aes(x = time, y = value)) +
  geom_hline(yintercept = var2) # tref is 10


encounter <- mizerEncounterTemp(params, n = params@initial_n, n_pp = params@initial_n_pp, n_other = params@initial_n_other, t = 15)

feeding_level <- mizerFeedingLevelTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       encounter = encounter, t = 15)

predRate <- mizerPredRateTemp(params, n = params@initial_n, 
                                       n_pp = params@initial_n_pp, n_other = params@initial_n_other, 
                                       feeding_level = feeding_level, t = 15)

```


Now writing a proper setup to enable temperature in mizer


```{r}
source("temperatureFuncs.R")
# a function that:
# adds the temperature parameters to the params object
# replace default rates in mizer with temperature mizer
# add temperature vector to param object

#' @paramTemp a dataframe with the temperature parameters to add to the mizer object 

mizerTemp <- function(temp_vec, params, paramTemp = NULL){
  
if(!is.null(paramTemp)){
params@species_params$Ea <- paramTemp$Ea
params@species_params$c_a <- paramTemp$c_a
params@species_params$Ed <- paramTemp$Ed
params@species_params$c_d <- paramTemp$c_d
params@species_params$t_d <- paramTemp$t_d
params@species_params$t_ref <- paramTemp$t_ref
}

other_params(params)$temperature <- temp_vec

params <- setRateFunction(params, "Encounter", "mizerEncounterTemp")
params <- setRateFunction(params, "FeedingLevel", "mizerFeedingLevelTemp")
params <- setRateFunction(params, "PredRate", "mizerPredRateTemp")

params <- setParams(params)
  
  return(params)
}


```


test


```{r}

params <- newTraitParams(no_sp = 4)
sim <- project(params)
plotBiomass(sim)

temperatureParams <- data.frame("Ea" = 0.63, "c_a" = 0, "Ed" = 6.3, "c_d" = 0,
                                "t_d" = 25, "t_ref" = 10)

temperature <- seq(10,30,length.out = 200)

paramsT <- mizerTemp(temp_vec = temperature, paramTemp = temperatureParams, params = params)

simT <- project(paramsT, t_max = length(temperature))

plotBiomass(simT)
```



Working :)


toothfish

Preferred temperature (Ref. 115969): 1.8 - 8.8, mean 4.5 °C (based on 1109 cells).


